---
title: "Ch. 12 Multilevel Models"
author: "A Solomon Kurz"
date: "`r format(Sys.Date())`"
output:
  github_document
---

```{r set-options_12, echo = FALSE, cache = FALSE}
options(width = 100)
```

# Multilevel Models

> Multilevel models... remember features of each cluster in the data as they learn about all of the clusters. Depending upon the variation among clusters, which is learned from the data as well, the model pools information across clusters. This pooling tends to improve estimates about each cluster. This improved estimation leads to several, more pragmatic sounding, benefits of the multilevel approach. (p. 356)

These benefits include:

* improved estimates for repeated sampling (i.e., in longitudinal data)
* improved estimates when there are imbalances among subsamples
* estimates of the variation across subsamples
* avoiding simplistic averaging by retaining variation across subsamples

> All of these benefits flow out of the same strategy and model structure. You learn one basic design and you get all of this for free.
>
> When it comes to regression, multilevel regression deserves to be the default approach. There are certainly contexts in which it would be better to use an old-fashioned single-level model. But the contexts in which multilevel models are superior are much more numerous. It is better to begin to build a multilevel analysis, and then realize it’s unnecessary, than to overlook it. And once you grasp the basic multilevel stragety, it becomes much easier to incorporate related tricks such as allowing for measurement error in the data and even model missing data itself (Chapter 14). (p. 356)

I’m totally on board with this. After learning about the multilevel model, I see it everywhere. For more on the sentiment it should be the default, check out McElreath’s blog post, [*Multilevel Regression as Default*](http://elevanth.org/blog/2017/08/24/multilevel-regression-as-default/).

## Example: Multilevel tadpoles

Let's get the `reedfrogs` data from rethinking.

```{r, message = F, warning = F}
library(rethinking)
data(reedfrogs)
d <- reedfrogs
```

Detach rethinking and load brms.

```{r, message = F}
rm(reedfrogs)
detach(package:rethinking, unload = T)
library(brms)
```

Go ahead and acquaint yourself with the `reedfrogs`.

```{r, message = F, warning = F}
library(tidyverse)

d %>%
  glimpse()
```

Making the `tank` cluster variable is easy.

```{r}
d <- 
  d %>%
  mutate(tank = 1:nrow(d))
```

Here's the formula for the un-pooled model in which each `tank` gets its own intercept.

\begin{align*}
\text{surv}_i        & \sim \text{Binomial} (n_i, p_i) \\
\text{logit} (p_i)   & = \alpha_{\text{tank}_i} \\
\alpha_{\text{tank}} & \sim \text{Normal} (0, 5)
\end{align*}

And $n_i = \text{density}_i$. Now we'll fit this simple aggregated binomial model much like we practiced in Chapter 10.

```{r b12.1, cache = T, message = F, warning = F}
b12.1 <- 
  brm(data = d, family = binomial,
      surv | trials(density) ~ 0 + factor(tank),
      prior(normal(0, 5), class = b),
      iter = 2000, warmup = 500, chains = 4, cores = 4,
      seed = 12)
```

The formula for the multilevel alternative is

\begin{align*}
\text{surv}_i        & \sim \text{Binomial} (n_i, p_i) \\
\text{logit} (p_i)   & = \alpha_{\text{tank}_i} \\
\alpha_{\text{tank}} & \sim \text{Normal} (\alpha, \sigma) \\
\alpha               & \sim \text{Normal} (0, 1) \\
\sigma               & \sim \text{HalfCauchy} (0, 1)
\end{align*}

You specify the corresponding multilevel model like this.

```{r b12.2, cache = T, message = F, warning = F}
b12.2 <- 
  brm(data = d, family = binomial,
      surv | trials(density) ~ 1 + (1 | tank),
      prior = c(prior(normal(0, 1), class = Intercept),
                prior(cauchy(0, 1), class = sd)),
      iter = 4000, warmup = 1000, chains = 4, cores = 4,
      seed = 12)
```

The syntax for the varying effects follows the [lme4 style](https://cran.r-project.org/web/packages/brms/vignettes/brms_overview.pdf), `( <varying parameter(s)> | <grouping variable(s)> )`. In this case `(1 | tank)` indicates only the intercept, `1`, varies by `tank`. The extent to which parameters vary is controlled by the prior, `prior(cauchy(0, 1), class = sd)`, which is <u>parameterized in the standard deviation metric</u>. Do note that last part. It's common in multilevel software to model in the variance metric, instead. 

Let's do the WAIC comparisons.

```{r}
b12.1 <- add_criterion(b12.1, "waic")
b12.2 <- add_criterion(b12.2, "waic")

w <- loo_compare(b12.1, b12.2, criterion = "waic")

print(w, simplify = F)
```

The `se_diff` is large relative to the `elpd_diff`. If we convert the $\text{elpd}$ difference to the WAIC metric, the message stays the same.

```{r}
cbind(waic_diff = w[, 1] * -2,
      se        = w[, 2] *  2)
```

I’m not going to show it here, but if you’d like a challenge, try comparing the models with the LOO. You’ll learn all about high `pareto_k` values, `kfold()` recommendations, and challenges implementing those `kfold()` recommendations. If you’re interested, pour yourself a calming adult  beverage, execute the code below, and check out the [*Kfold(): “Error: New factor levels are not allowed”*](https://discourse.mc-stan.org/t/kfold-error-new-factor-levels-are-not-allowed/8299) thread in the Stan forums.

```{r, eval = F}
b12.1 <- add_criterion(b12.1, "loo")
b12.2 <- add_criterion(b12.2, "loo")
```

But back on track, here's our prep work for Figure 12.1.

```{r}
post <- posterior_samples(b12.2, add_chain = T)

post_mdn <- 
  coef(b12.2, robust = T)$tank[, , ] %>% 
  as_tibble() %>% 
  bind_cols(d) %>%
  mutate(post_mdn = inv_logit_scaled(Estimate))

post_mdn
```

For kicks and giggles, let's use a [FiveThirtyEight-like theme](https://github.com/alex23lemm/theme_fivethirtyeight) for this chapter’s plots. An easy way to do so is with help from the [ggthemes package](https://cran.r-project.org/web/packages/ggthemes/index.html).

```{r}
# install.packages("ggthemes", dependencies = T) 

library(ggthemes) 
``` 

Finally, here's the ggplot2 code to reproduce Figure 12.1.

```{r, fig.width = 5, fig.height = 4}
post_mdn %>%
  ggplot(aes(x = tank)) +
  geom_hline(yintercept = inv_logit_scaled(median(post$b_Intercept)), linetype = 2, size = 1/4) +
  geom_vline(xintercept = c(16.5, 32.5), size = 1/4) +
  geom_point(aes(y = propsurv), color = "orange2") +
  geom_point(aes(y = post_mdn), shape = 1) +
  coord_cartesian(ylim = c(0, 1)) +
  scale_x_continuous(breaks = c(1, 16, 32, 48)) +
  labs(title    = "Multilevel shrinkage!",
       subtitle = "The empirical proportions are in orange while the model-\nimplied proportions are the black circles. The dashed line is\nthe model-implied average survival proportion.") +
  annotate("text", x = c(8, 16 + 8, 32 + 8), y = 0, 
           label = c("small tanks", "medium tanks", "large tanks")) +
  theme_fivethirtyeight() +
  theme(panel.grid = element_blank())
```

Here is our version of Figure 12.2.a. 

```{r, fig.width = 3, fig.height = 3}
# this makes the output of `sample_n()` reproducible
set.seed(12)

post %>% 
  sample_n(100) %>% 
  expand(nesting(iter, b_Intercept, sd_tank__Intercept),
         x = seq(from = -4, to = 5, length.out = 100)) %>% 

  ggplot(aes(x = x, group = iter)) +
  geom_line(aes(y = dnorm(x, b_Intercept, sd_tank__Intercept)),
            alpha = .2, color = "orange2") +
  labs(title = "Population survival distribution",
       subtitle = "The Gaussians are on the log-odds scale.") +
  scale_y_continuous(NULL, breaks = NULL) +
  coord_cartesian(xlim = c(-3, 4)) + 
  theme_fivethirtyeight() +
  theme(plot.title    = element_text(size = 13),
        plot.subtitle = element_text(size = 10))
```

Note the uncertainty in terms of both location $\alpha$ and scale $\sigma$. Now here's the code for Figure 12.2.b.

```{r, fig.width = 3, fig.height = 3}
ggplot(data = post, 
       aes(x = rnorm(n    = nrow(post), 
                     mean = b_Intercept, 
                     sd   = sd_tank__Intercept) %>% 
             inv_logit_scaled())) +
  geom_density(size = 0, fill = "orange2") +
  scale_y_continuous(NULL, breaks = NULL) +
  ggtitle("Probability of survival") +
  theme_fivethirtyeight()
```

Note how we sampled 12,000 imaginary `tanks` rather than McElreath's 8,000. This is because we had 12,000 HMC iterations (i.e., execute `nrow(post)`).

The `aes()` code, above, was a bit much. To get a sense of how it worked, consider this:

```{r}
set.seed(12)

rnorm(n    = 1, 
      mean = post$b_Intercept, 
      sd   = post$sd_tank__Intercept) %>% 
  inv_logit_scaled()
```

First, we took one random draw from a normal distribution with a mean of the first row in `post$b_Intercept` and a standard deviation of the value from the first row in `post$sd_tank__Intercept`, and passed it through the `inv_logit_scaled()` function. By replacing the `1` with `nrow(post)`, we do this `nrow(post)` times (i.e., 12,000). Our orange density, then, is the summary of that process.

#### Overthinking: Prior for variance components.

Yep, you can use the exponential distribution for your priors in brms. Here it is for model `b12.2`.

```{r b12.2.e, cache = T, message = F, results = "hide"}
b12.2.e <- 
  update(b12.2,
         prior = c(prior(normal(0, 1), class = Intercept),
                   prior(exponential(1), class = sd)))
```

The model summary:

```{r}
print(b12.2.e)
```

If you're curious how the exponential prior compares to the posterior, you might just plot.

```{r, fig.width = 3.75, fig.height = 3.25}
tibble(x = seq(from = 0, to = 6, by = .01)) %>% 
  
  ggplot() +
  # the prior
  geom_ribbon(aes(x = x, ymin = 0, ymax = dexp(x, rate = 1)),
              fill = "orange2", alpha = 1/3) +
  # the posterior
  geom_density(data = posterior_samples(b12.2.e),
               aes(x = sd_tank__Intercept), 
               fill = "orange2", size = 0) +
  scale_y_continuous(NULL, breaks = NULL) +
  coord_cartesian(xlim = c(0, 5)) +
  labs(title    = "Bonus prior/posterior plot\nfor sd_tank__Intercept",
       subtitle = "The prior is the semitransparent ramp in the\nbackground. The posterior is the solid orange\nmound.") +
  theme_fivethirtyeight()
```  

## Varying effects and the underfitting/overfitting trade-off

> Varying intercepts are just regularized estimates, but adaptively regularized by estimating how diverse the clusters are while estimating the features of each cluster. This fact is not easy to grasp…
>
> A major benefit of using varying effects estimates, instead of the empirical raw estimates, is that they provide more accurate estimates of the individual cluster (tank) intercepts. On average, the varying effects actually provide a better estimate of the individual tank (cluster) means. The reason that the varying intercepts provides better estimates is that they do a better job trading off underfitting and overfitting. (p. 364)

In this section, we explicate this by contrasting three perspectives:

* Complete pooling (i.e., a single-$\alpha$ model)
* No pooling (i.e., the single-level $\alpha_{\text{tank}_i}$ model)
* Partial pooling (i.e., the multilevel model for which $\alpha_{\text{tank}} \sim \text{Normal} (\alpha, \sigma)$)

> To demonstrate [the magic of the multilevel model], we’ll simulate some tadpole data. That way, we’ll know the true per-pond survival probabilities. Then we can compare the no-pooling estimates to the partial pooling estimates, by computing how close each gets to the true values they are trying to estimate. The rest of this section shows how to do such a simulation. (p. 365)

### The model.

The simulation formula should look familiar.

\begin{align*}
\text{surv}_i        & \sim \text{Binomial} (n_i, p_i) \\
\text{logit} (p_i)   & = \alpha_{\text{pond}_i} \\
\alpha_{\text{pond}} & \sim \text{Normal} (\alpha, \sigma) \\
\alpha               & \sim \text{Normal} (0, 1) \\
\sigma               & \sim \text{HalfCauchy} (0, 1)
\end{align*}

### Assign values to the parameters.

```{r}
a       <-  1.4
sigma   <-  1.5
n_ponds <- 60

set.seed(12)
(
  dsim <- 
  tibble(pond   = 1:n_ponds,
         ni     = rep(c(5, 10, 25, 35), each = n_ponds / 4) %>% as.integer(),
         true_a = rnorm(n = n_ponds, mean = a, sd = sigma))
  )
```

### Sumulate survivors.

> Each pond $i$ has $n_i$ potential survivors, and nature flips each tadpole’s coin, so to speak, with probability of survival $p_i$. This probability $p_i$ is implied by the model definition, and is equal to:
>
> $$p_i = \frac{\text{exp} (\alpha_i)}{1 + \text{exp} (\alpha_i)}$$
>
> The model uses a logit link, and so the probability is defined by the [`inv_logit_scaled()`] function. (p. 367)

```{r}
set.seed(12)
(
  dsim <-
  dsim %>%
  mutate(si = rbinom(n = n(), prob = inv_logit_scaled(true_a), size = ni))
  )
```

### Compute the no-pooling estimates.

The no-pooling estimates (i.e., $\alpha_{\text{tank}_i}$) are the results of simple algebra.

```{r}
(
  dsim <-
  dsim %>%
  mutate(p_nopool = si / ni)
  )
```

"These are the same no-pooling estimates you’d get by fitting a model with a dummy variable for each pond and flat priors that induce no regularization" (p. 367).

### Compute the partial-pooling estimates.

To follow along with McElreath, set `chains = 1, cores = 1` to fit with one chain.

```{r b12.3, cache = T, message = F, warning = F, results = "hide"}
b12.3 <- 
  brm(data = dsim, family = binomial,
      si | trials(ni) ~ 1 + (1 | pond),
      prior = c(prior(normal(0, 1), class = Intercept),
                prior(cauchy(0, 1), class = sd)),
      iter = 10000, warmup = 1000, chains = 1, cores = 1,
      seed = 12)
```

```{r}
print(b12.3)
```

I'm not aware that you can use McElreath's `depth=2` trick in brms for `summary()` or `print()`. But can get that information with the `coef()` function. 

```{r}
coef(b12.3)$pond[c(1:2, 59:60), , ] %>% 
  round(digits = 2)
```

Note how we just peeked at the top and bottom two rows with the `c(1:2, 59:60)` part of the code, there. Somewhat discouragingly, `coef()` doesn’t return the 'Eff.Sample' or 'Rhat' columns as in McElreath’s output. We can still extract that information, though. For $\hat{R}$, the solution is simple; use the `brms::rhat()` function.

```{r}
rhat(b12.3)
```

Extracting the 'Eff.Sample' values is a little more complicated. There is no `effsamples()` function. However, we do have `neff_ratio()`. 

```{r}
neff_ratio(b12.3)
```

The `brms::neff_ratio()` function returns ratios of the effective samples over the total number of post-warmup iterations. So if we know the `neff_ratio()` values and the number of post-warmup iterations, the 'Eff.Sample' values are just a little algebra away. A quick solution is to look at the 'total post-warmup samples' line at the top of our `print()` output. Another way is to extract that information from our `brm()` fit object. I’m not aware of a way to do that directly, but we can extract the `iter` value (i.e., `b12.2$fit@sim$iter`), the `warmup` value (i.e., `b12.2$fit@sim$warmup`), and the number of `chains` (i.e., `b12.2$fit@sim$chains`). With those values in hand, simple algebra will return the 'total post-warmup samples' value. E.g.,

```{r}
(n_iter <- (b12.3$fit@sim$iter - b12.3$fit@sim$warmup) * b12.3$fit@sim$chains)
```

And now we have `n_iter`, we can calculate the 'Eff.Sample' values.

```{r}
neff_ratio(b12.3) %>% 
  data.frame() %>% 
  rownames_to_column() %>% 
  set_names("parameter", "neff_ratio") %>% 
  mutate(eff_sample = (neff_ratio * n_iter) %>% round(digits = 0)) %>% 
  head()
```

Digressions aside, let's get ready for the diagnostic plot of Figure 12.3.

```{r}
dsim %>% 
  glimpse()
```

```{r}
# we could have included this step in the block of code below, if we wanted to
p_partpool <- 
  coef(b12.3)$pond[, , ] %>% 
  as_tibble() %>%
  transmute(p_partpool = inv_logit_scaled(Estimate))

dsim <- 
  dsim %>%
  bind_cols(p_partpool) %>% 
  mutate(p_true         = inv_logit_scaled(true_a)) %>%
  mutate(nopool_error   = abs(p_nopool   - p_true),
         partpool_error = abs(p_partpool - p_true))

dsim %>% 
  glimpse()
```

Here is our code for Figure 12.3. The extra data processing for `dfline` is how we get the values necessary for the horizontal summary lines.

```{r, fig.width = 5, fig.height = 5}
dfline <- 
  dsim %>%
  select(ni, nopool_error:partpool_error) %>%
  gather(key, value, -ni) %>%
  group_by(key, ni) %>%
  summarise(mean_error = mean(value)) %>%
  mutate(x    = c( 1, 16, 31, 46),
         xend = c(15, 30, 45, 60))
  
dsim %>% 
  ggplot(aes(x = pond)) +
  geom_vline(xintercept = c(15.5, 30.5, 45.4), 
             color = "white", size = 2/3) +
  geom_point(aes(y = nopool_error), color = "orange2") +
  geom_point(aes(y = partpool_error), shape = 1) +
  geom_segment(data = dfline, 
               aes(x = x, xend = xend, 
                   y = mean_error, yend = mean_error),
               color = rep(c("orange2", "black"), each = 4),
               linetype = rep(1:2, each = 4)) +
  scale_x_continuous(breaks = c(1, 10, 20, 30, 40, 50, 60)) +
  annotate("text", x = c(15 - 7.5, 30 - 7.5, 45 - 7.5, 60 - 7.5), y = .45, 
           label = c("tiny (5)", "small (10)", "medium (25)", "large (35)")) +
  labs(y        = "absolute error",
       title    = "Estimate error by model type",
       subtitle = "The horizontal axis displays pond number. The vertical axis measures\nthe absolute error in the predicted proportion of survivors, compared to\nthe true value used in the simulation. The higher the point, the worse\nthe estimate. No-pooling shown in orange. Partial pooling shown in black.\nThe orange and dashed black lines show the average error for each kind\nof estimate, across each initial density of tadpoles (pond size). Smaller\nponds produce more error, but the partial pooling estimates are better\non average, especially in smaller ponds.") +
  theme_fivethirtyeight() +
  theme(panel.grid    = element_blank(),
        plot.subtitle = element_text(size = 10))
```

If you wanted to quantify the difference in simple summaries, you might do something like this:

```{r}
dsim %>%
  select(ni, nopool_error:partpool_error) %>%
  gather(key, value, -ni) %>%
  group_by(key) %>%
  summarise(mean_error   = mean(value) %>% round(digits = 3),
            median_error = median(value) %>% round(digits = 3))
```

I originally learned about the multilevel in order to work with [longitudinal data](http://gseacademic.harvard.edu/alda/). In that context, I found the basic principles of a multilevel structure quite intuitive. The concept of partial pooling, however, took me some time to wrap my head around. If you’re struggling with this, be patient and keep chipping away. 

When McElreath [lectured on this topic in 2015](https://www.youtube.com/watch?v=82TaniPgzQc&t=2048s&frags=pl%2Cwn), he traced partial pooling to statistician [Charles M. Stein](http://bulletin.imstat.org/2017/05/obituary-charles-m-stein-1920-2016/). In 1977, Efron and Morris wrote the now classic paper, [*Stein’s Paradox in Statistics*](http://statweb.stanford.edu/%7Eckirby/brad/other/Article1977.pdf), which does a nice job breaking down why partial pooling can be so powerful. One of the primary examples they used in the paper was of 1970 batting average data. If you’d like more practice seeing how partial pooling works--or if you just like baseball--, check out my blog post, [*Stein’s Paradox and What Partial Pooling Can Do For You*](https://solomonkurz.netlify.com/post/stein-s-paradox-and-what-partial-pooling-can-do-for-you/).

#### Overthinking: Repeating the pond simulation. 

Within the brms workflow, we can reuse a compiled model with `update()`. But first, we'll simulate new data.

```{r}
a       <-  1.4
sigma   <-  1.5
n_ponds <- 60

set.seed(1999)  # for new data, set a new seed
new_dsim <- 
  tibble(pond   = 1:n_ponds,
         ni     = rep(c(5, 10, 25, 35), each = n_ponds / 4) %>% as.integer(),
         true_a = rnorm(n = n_ponds, mean = a, sd = sigma)) %>% 
    mutate(si       = rbinom(n = n(), prob = inv_logit_scaled(true_a), size = ni)) %>% 
    mutate(p_nopool = si / ni)

glimpse(new_dsim)
```

Fit the new model.

```{r b12.3_new, cache = T, message = F, warning = F, results = "hide"}
b12.3_new <- 
  update(b12.3,
         newdata = new_dsim,
         iter = 10000, warmup = 1000, chains = 1, cores = 1)
```

```{r}
print(b12.3_new)
```

Why not plot the first simulation versus the second one?

```{r, fig.width = 8, fig.height = 4.5}
bind_rows(posterior_samples(b12.3),
          posterior_samples(b12.3_new)) %>%
  mutate(model = rep(c("b12.3", "b12.3_new"), each = n() / 2)) %>% 

  ggplot(aes(x = b_Intercept, y = sd_pond__Intercept)) +
  stat_density_2d(geom = "raster", 
                  aes(fill = stat(density)), 
                  contour = F, n = 200) +
  geom_vline(xintercept = a,     color = "orange3", linetype = 3) +
  geom_hline(yintercept = sigma, color = "orange3", linetype = 3) +
  scale_fill_gradient(low = "grey25", high = "orange3") +
  ggtitle("Our simulation posteriors contrast a bit",
          subtitle = expression(paste(alpha, " is on the x and ", sigma, " is on the y, both in log-odds. The dotted lines intersect at the true values."))) +
  coord_cartesian(xlim = c(.7, 2),
                  ylim = c(.8, 1.9)) +
  theme_fivethirtyeight() +
  theme(legend.position = "none",
        panel.grid      = element_blank()) +
  facet_wrap(~model, ncol = 2)
```

If you'd like the `stanfit` portion of your `brm()` object, subset with `$fit`. Take `b12.3`, for example. You might check out its structure via `b12.3$fit %>% str()`. Here's the actual Stan code.

```{r}
b12.3$fit@ stanmodel
```

And you can get the data of a given `brm()` fit object like so.

```{r}
b12.3$data %>% 
  head()
```

## More than one type of cluster

"We can use and often should use more than one type of cluster in the same model" (p. 370).

### Multilevel chimpanzees.

The initial multilevel update from model `b10.4` from the last chapter follows the statistical formula

\begin{align*}
\text{left_pull}_i   & \sim \text{Binomial} (n_i = 1, p_i) \\
\text{logit} (p_i)    & = \alpha + \alpha_{\text{actor}_i} + (\beta_1 + \beta_2 \text{condition}_i) \text{prosoc_left}_i \\
\alpha_{\text{actor}} & \sim \text{Normal} (0, \sigma_{\text{actor}}) \\
\alpha                & \sim \text{Normal} (0, 10) \\
\beta_1               & \sim \text{Normal} (0, 10) \\
\beta_2               & \sim \text{Normal} (0, 10) \\
\sigma_{\text{actor}} & \sim \text{HalfCauchy} (0, 1)
\end{align*}

> Notice that $\alpha$ is inside the linear model, not inside the Gaussian prior for $\alpha_\text{actor}$. This is mathematically equivalent to what [we] did with the tadpoles earlier in the chapter. You can always take the mean out of a Gaussian distribution and treat that distribution as a constant plus a Gaussian distribution centered on zero.
>
> This might seem a little weird at first, so it might help train your intuition by experimenting in R. (p. 371)

Behold our two identical Gaussians in a tidy tibble.

```{r}
set.seed(12)
two_gaussians <- 
  tibble(y1 =      rnorm(n = 1e4, mean = 10, sd = 1),
         y2 = 10 + rnorm(n = 1e4, mean = 0,  sd = 1))
```

Let's follow McElreath's advice to make sure they are same by superimposing the density of one on the other.

```{r, fig.width = 3.5, fig.height = 3}
two_gaussians %>%
  
  ggplot() +
  geom_density(aes(x = y1), 
               size = 0, fill = "orange1", alpha = 1/3) +
  geom_density(aes(x = y2), 
               size = 0, fill = "orange4", alpha = 1/3) +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(title = "Our simulated Gaussians") +
  theme_fivethirtyeight()
```

Yep, those Gaussians look about the same.

Let's get the `chimpanzees` data from rethinking.

```{r, message = F}
library(rethinking)
data(chimpanzees)
d <- chimpanzees
```

Detach rethinking and reload brms.

```{r, message = F}
rm(chimpanzees)
detach(package:rethinking, unload = T)
library(brms)
```

For our brms model with varying intercepts for `actor` but not `block`, we employ the `pulled_left ~ 1 + ... + (1 | actor)` syntax, specifically omitting a `(1 | block)` section.

```{r b12.4, cache = T, message = F, warning = F}
b12.4 <- 
  brm(data = d, family = binomial,
      pulled_left | trials(1) ~ 1 + prosoc_left + prosoc_left:condition + (1 | actor),
      prior = c(prior(normal(0, 10), class = Intercept),
                prior(normal(0, 10), class = b),
                prior(cauchy(0, 1), class = sd)),
      # I'm using 4 cores, instead of the `cores=3` in McElreath's code
      iter = 5000, warmup = 1000, chains = 4, cores = 4,  
      control = list(adapt_delta = 0.95),
      seed = 12)
```

The initial solutions came with a few divergent transitions. Increasing `adapt_delta` to `0.95` solved the problem. You can also solve the problem with more strongly regularizing priors such as `normal(0, 2)` on the intercept and slope parameters (see [recommendations from the Stan team](https://github.com/stan-dev/stan/wiki/Prior-Choice-Recommendations)). Consider trying both methods and comparing the results. They're similar. 

Here we add the `actor`-level deviations to the fixed intercept, the grand mean.

```{r}
post <- posterior_samples(b12.4)

post %>%
  select(starts_with("r_actor")) %>% 
  gather() %>%
  # this is how we might add the grand mean to the actor-level deviations
  mutate(value = value + post$b_Intercept) %>% 
  group_by(key) %>%
  summarise(mean = mean(value) %>% round(digits = 2))
```

Here's another way to get at the same information, this time using `coef()` and a little formatting help from the `stringr::str_c()` function. Just for kicks, we'll throw in the 95% intervals, too.

```{r, message = F}
coef(b12.4)$actor[, c(1, 3:4), 1] %>%
  as_tibble() %>% 
  round(digits = 2) %>%
  # here we put the credible intervals in an APA-6-style format
  mutate(`95% CIs` = str_c("[", Q2.5, ", ", Q97.5, "]"),
         actor     = str_c("chimp #", 1:7)) %>%
  rename(mean = Estimate) %>%
  select(actor, mean, `95% CIs`) %>% 
  knitr::kable()
```

If you prefer the posterior median to the mean, just add a `robust = T` argument inside the `coef()` function.

### Two types of cluster.

The full statistical model follows the form

\begin{align*}
\text{left_pull}_i    & \sim \text{Binomial} (n_i = 1, p_i) \\
\text{logit} (p_i)    & = \alpha + \alpha_{\text{actor}_i} + \alpha_{\text{block}_i} + (\beta_1 + \beta_2 \text{condition}_i) \text{prosoc_left}_i \\
\alpha_{\text{actor}} & \sim \text{Normal} (0, \sigma_{\text{actor}}) \\
\alpha_{\text{block}} & \sim \text{Normal} (0, \sigma_{\text{actor}}) \\
\alpha                & \sim \text{Normal} (0, 10) \\
\beta_1               & \sim \text{Normal} (0, 10) \\
\beta_2               & \sim \text{Normal} (0, 10) \\
\sigma_{\text{actor}} & \sim \text{HalfCauchy} (0, 1) \\
\sigma_{\text{block}} & \sim \text{HalfCauchy} (0, 1)
\end{align*}

Our brms model with varying intercepts for both `actor` and `block` now employs the `... (1 | actor) + (1 | block)` syntax.

```{r b12.5, cache = T, message = F, warning = F}
b12.5 <- 
  update(b12.4,
         newdata = d,
         formula = pulled_left | trials(1) ~ 1 + prosoc_left + prosoc_left:condition + 
           (1 | actor) + (1 | block),
         iter = 6000, warmup = 1000, cores = 4, chains = 4, 
         control = list(adapt_delta = 0.99),
         seed = 12)
```

This time we increased `adapt_delta` to `0.99` to avoid divergent transitions. We can look at the primary coefficients with `print()`. McElreath encouraged us to inspect the trace plots. Here they are.

```{r, message = F, warning = F, fig.width = 8, fig.height = 7}
library(bayesplot)
color_scheme_set("orange")

post <- posterior_samples(b12.5, add_chain = T)

post %>% 
  select(-lp__, -iter) %>% 
  mcmc_trace(facet_args = list(ncol = 4)) +
  scale_x_continuous(breaks = c(0, 2500, 5000)) +
  theme_fivethirtyeight() +
  theme(legend.position = c(.75, .06))
```

The trace plots look great. We may as well examine the $n_\text{eff} / N$ ratios, too.

```{r, fig.width = 8, fig.height = 4}
neff_ratio(b12.5) %>% 
  mcmc_neff() +
  theme_fivethirtyeight()
```

About half of them are lower than we might like, but none are in the embarrassing $n_\text{eff} / N \leq .1$ range. Let's look at the summary of the main parameters.

```{r}
print(b12.5)
```

This time, we'll need to use `brms::ranef()` to get those `depth=2`-type estimates in the same metric displayed in the text. With `ranef()`, you get the group-specific estimates in a deviance metric. The `coef()` function, in contrast, yields the group-specific estimates in what you might call the natural metric. We'll get more language for this in the next chapter.

```{r}
ranef(b12.5)$actor[, , "Intercept"] %>% 
  round(digits = 2)

ranef(b12.5)$block[, , "Intercept"] %>% 
  round(digits = 2)
```

We might make the coefficient plot of Figure 12.4.a like this:

```{r, fig.width = 3.5, fig.height = 3.5}
stanplot(b12.5, pars = c("^r_", "^b_", "^sd_")) +
  theme_fivethirtyeight() +
  theme(axis.text.y = element_text(hjust = 0))
```

Once we get the posterior samples, it's easy to compare the random variances as in Figure 12.4.b.

```{r, fig.width = 3, fig.height = 3}
post %>%
  ggplot(aes(x = sd_actor__Intercept)) +
  geom_density(size = 0, fill = "orange1", alpha = 3/4) +
  geom_density(aes(x = sd_block__Intercept), 
               size = 0, fill = "orange4", alpha = 3/4)  +
  scale_y_continuous(NULL, breaks = NULL) +
  coord_cartesian(xlim = c(0, 4)) +
  labs(title = expression(sigma)) +
  annotate("text", x = 2/3, y = 2, label = "block", color = "orange4") +
  annotate("text", x = 2, y = 3/4, label = "actor", color = "orange1") +
  theme_fivethirtyeight()
```

We might compare our models by their PSIS-LOO values.

```{r}
b12.4 <- add_criterion(b12.4, "loo")
b12.5 <- add_criterion(b12.5, "loo")

loo_compare(b12.4, b12.5) %>% 
  print(simplify = F)
```

The two models yield nearly-equivalent information criteria values. Yet recall what McElreath wrote: "There is nothing to gain here by selecting either model. The comparison of the two models tells a richer story" (p. 367).

## Multilevel posterior predictions

> … producing implied predictions from a fit model, is very helpful for understanding what the model means. Every model is a merger of sense and nonsense. When we understand a model, we can find its sense and control its nonsense. But as models get more complex, it is very difficult to impossible to understand them just by inspecting tables of posterior means and intervals. Exploring implied posterior predictions helps much more…
>
> … The introduction of varying effects does introduce nuance, however.
>
> First, we should no longer expect the model to exactly retrodict the sample, because adaptive regularization has as its goal to trade off poorer fit in sample for better inference and hopefully better fit out of sample. This is what shrinkage does for us…
>
Second, "prediction" in the context of a multilevel model requires additional choices. If we wish to validate a model against the specific clusters used to fit the model, that is one thing. But if we instead wish to compute predictions for new clusters, other than the one observed in the sample, that is quite another. We’ll consider each of these in turn, continuing to use the chimpanzees model from the previous section. (p. 376)

### Posterior prediction for same clusters.

Like McElreath did in the text, we'll do this two ways. Recall we use `brms::fitted()` in place of `rethinking::link()`.

```{r, message = F, warning = F}
chimp <- 2
nd <-
  tibble(prosoc_left = c(0, 1, 0, 1),
         condition   = c(0, 0, 1, 1),
         actor       = chimp)

(
  chimp_2_fitted <-
  fitted(b12.4,
         newdata = nd) %>% 
  as_tibble() %>% 
  mutate(condition = factor(c("0/0", "1/0", "0/1", "1/1"), 
                            levels = c("0/0", "1/0", "0/1", "1/1")))
  )
```

```{r}
(
  chimp_2_d <-
  d %>% 
  filter(actor == chimp) %>% 
  group_by(prosoc_left, condition) %>% 
  summarise(prob = mean(pulled_left)) %>% 
  ungroup() %>% 
  mutate(condition = str_c(prosoc_left, "/", condition)) %>% 
  mutate(condition = factor(condition, levels = c("0/0", "1/0", "0/1", "1/1")))
)
```

McElreath didn't show the corresponding plot in the text. It might look like this.

```{r, fig.width = 2.5, fig.height = 4}
chimp_2_fitted %>%
  # if you want to use `geom_line()` or `geom_ribbon()` with a factor on the x axis,
  # you need to code something like `group = 1` in `aes()`
  ggplot(aes(x = condition, y = Estimate, group = 1)) +
  geom_ribbon(aes(ymin = Q2.5, ymax = Q97.5), fill = "orange1") +
  geom_line(color = "blue") +
  geom_point(data = chimp_2_d,
             aes(y = prob),
             color = "grey25") +
  ggtitle("Chimp #2",
          subtitle = "The posterior mean and 95%\nintervals are the blue line\nand orange band, respectively.\nThe empirical means are\nthe charcoal dots.") +
  coord_cartesian(ylim = c(.75, 1)) +
  theme_fivethirtyeight() +
  theme(plot.subtitle = element_text(size = 10))
```

Do note how severely we’ve restricted the y-axis range. But okay, now let's do things by hand. We'll need to extract the posterior samples and look at the structure of the data.

```{r}
post <- posterior_samples(b12.4)

glimpse(post)
```

McElreath didn't show what his R code 12.29 `dens( post$a_actor[,5] )` would look like. But here's our analogue.

```{r, fig.width = 4, fig.height = 2.5}
post %>%
  transmute(actor_5 = `r_actor[5,Intercept]`) %>% 
  
  ggplot(aes(x = actor_5)) +
  geom_density(size = 0, fill = "blue") +
  scale_y_continuous(breaks = NULL) +
  ggtitle("Chimp #5's density") +
  theme_fivethirtyeight()
```

And because we made the density only using the `r_actor[5,Intercept]` values (i.e., we didn’t add `b_Intercept` to them), the density is in a deviance-score metric.

McElreath built his own `link()` function. Here we'll build an alternative to `fitted()`.

```{r}
# our hand-made `brms::fitted()` alternative
my_fitted <- function(prosoc_left, condition){
  post %>%
    transmute(fitted = (b_Intercept + 
                          `r_actor[5,Intercept]` + 
                          b_prosoc_left * prosoc_left + 
                          `b_prosoc_left:condition` * prosoc_left * condition) %>% 
                inv_logit_scaled())
}

# the posterior summaries
(
  chimp_5_my_fitted <-
  tibble(prosoc_left = c(0, 1, 0, 1),
       condition     = c(0, 0, 1, 1)) %>% 
  mutate(post = map2(prosoc_left, condition, my_fitted)) %>% 
  unnest() %>% 
  mutate(condition = str_c(prosoc_left, "/", condition)) %>% 
  mutate(condition = factor(condition, levels = c("0/0", "1/0", "0/1", "1/1"))) %>% 
  group_by(condition) %>% 
  tidybayes::mean_qi(fitted)
  )

# the empirical summaries
chimp <- 5
(
  chimp_5_d <-
  d %>% 
  filter(actor == chimp) %>% 
  group_by(prosoc_left, condition) %>% 
  summarise(prob = mean(pulled_left)) %>% 
  ungroup() %>% 
  mutate(condition = str_c(prosoc_left, "/", condition)) %>% 
  mutate(condition = factor(condition, levels = c("0/0", "1/0", "0/1", "1/1")))
)
```

Okay, let's see how good we are at retrodicting the `pulled_left` probabilities for `actor == 5`.

```{r, fig.width = 2.5, fig.height = 3.75}
chimp_5_my_fitted %>%
  ggplot(aes(x = condition, y = fitted, group = 1)) +
  geom_ribbon(aes(ymin = .lower, ymax = .upper), fill = "orange1") +
  geom_line(color = "blue") +
  geom_point(data = chimp_5_d,
             aes(y = prob),
             color = "grey25") +
  ggtitle("Chimp #5",
          subtitle = "This plot is like the last except\nwe did more by hand.") +
  coord_cartesian(ylim = 0:1) +
  theme_fivethirtyeight() +
  theme(plot.subtitle = element_text(size = 10))
```

Not bad.

### Posterior prediction for new clusters. 

By average actor, McElreath referred to a chimp with an intercept exactly at the population mean $\alpha$. So this time we’ll only be working with the population parameters, or what are also sometimes called the fixed effects. When using `brms::posterior_samples()` output, this would mean working with columns beginning with the `b_` prefix (i.e., `b_Intercept`, `b_prosoc_left`, and `b_prosoc_left:condition`).

```{r}
post_average_actor <-
  post %>% 
  # here we use the linear regression formula to get the log_odds for the 4 conditions
  transmute(`0/0` = b_Intercept,
            `1/0` = b_Intercept + b_prosoc_left,
            `0/1` = b_Intercept,
            `1/1` = b_Intercept + b_prosoc_left + `b_prosoc_left:condition`) %>%
  # with `mutate_all()` we can convert the estimates to probabilities in one fell swoop
  mutate_all(inv_logit_scaled) %>% 
  # putting the data in the long format and grouping by condition (i.e., `key`)
  gather() %>%
  mutate(key = factor(key, level = c("0/0", "1/0", "0/1", "1/1"))) %>% 
  group_by(key) %>%
  # here we get the summary values for the plot
  summarise(m  = mean(value),
            # note we're using 80% intervals
            ll = quantile(value, probs = .1),
            ul = quantile(value, probs = .9))

post_average_actor
```

Figure 12.5.a.

```{r, fig.width = 2.5, fig.height = 3.25}
p1 <-
  post_average_actor %>%
  ggplot(aes(x = key, y = m, group = 1)) +
  geom_ribbon(aes(ymin = ll, ymax = ul), fill = "orange1") +
  geom_line(color = "blue") +
  ggtitle("Average actor") +
  coord_cartesian(ylim = 0:1) +
  theme_fivethirtyeight() +
  theme(plot.title = element_text(size = 14, hjust = .5))

p1
```

If we want to depict the variability across the chimps, we need to include `sd_actor__Intercept` into the calculations. In the first block of code, below, we simulate a bundle of new intercepts defined by

$$\alpha_\text{actor} \sim \text{Normal} (0, \sigma_\text{actor})$$

```{r}
# the random effects
set.seed(12.42)
ran_ef <-
  tibble(random_effect = rnorm(n = 1000, mean = 0, sd = post$sd_actor__Intercept)) %>% 
  # with the `., ., ., .` syntax, we quadruple the previous line 
  bind_rows(., ., ., .) 

# the fixed effects (i.e., the population parameters)
fix_ef <-
  post %>% 
  slice(1:1000) %>%
  transmute(`0/0` = b_Intercept,
            `1/0` = b_Intercept + b_prosoc_left,
            `0/1` = b_Intercept,
            `1/1` = b_Intercept + b_prosoc_left + `b_prosoc_left:condition`) %>%
  gather() %>%
  rename(condition    = key, 
         fixed_effect = value) %>% 
  mutate(condition = factor(condition, level = c("0/0", "1/0", "0/1", "1/1")))

# combine them
ran_and_fix_ef <-
  bind_cols(ran_ef, fix_ef) %>%
  mutate(intercept = fixed_effect + random_effect) %>%
  mutate(prob      = inv_logit_scaled(intercept))

# to simplify things, we'll reduce them to summaries
(
  marginal_effects <-
  ran_and_fix_ef %>%
  group_by(condition) %>%
  summarise(m  = mean(prob),
            ll = quantile(prob, probs = .1),
            ul = quantile(prob, probs = .9))
  )
```

Behold Figure 12.5.b.

```{r, fig.width = 2.5, fig.height = 3.25}
p2 <-
  marginal_effects %>%
  ggplot(aes(x = condition, y = m, group = 1)) +
  geom_ribbon(aes(ymin = ll, ymax = ul), fill = "orange1") +
  geom_line(color = "blue") +
  ggtitle("Marginal of actor") +
  coord_cartesian(ylim = 0:1) +
  theme_fivethirtyeight() +
  theme(plot.title = element_text(size = 14, hjust = .5))

p2
```

Figure 12.5.c just takes a tiny bit more wrangling. 

```{r, fig.width = 2.5, fig.height = 3.25}
p3 <-
  ran_and_fix_ef %>%
  mutate(iter = rep(1:1000, times = 4)) %>%
  filter(iter %in% c(1:50)) %>%
  
  ggplot(aes(x = condition, y = prob, group = iter)) +
  theme_fivethirtyeight() +
  ggtitle("50 simulated actors") +
  coord_cartesian(ylim = 0:1) +
  geom_line(alpha = 1/2, color = "orange3") +
  theme(plot.title = element_text(size = 14, hjust = .5))

p3
```

For the finale, we'll stitch the three plots together.

```{r, fig.width = 7.5, fig.height = 3.25, message = F, warning = F}
library(gridExtra)

grid.arrange(p1, p2, p3, ncol = 3)
```

#### Bonus: Let's use `fitted()` this time. 

We just made those plots using various wrangled versions of `post`, the data frame returned by `posterior_samples(b.12.4)`. If you followed along closely, part of what made that a great exercise is that it forced you to consider what the various vectors in `post` meant with respect to the model formula. But it’s also handy to see how to do that from a different perspective. So in this section, we’ll repeat that process by relying on the `fitted()` function, instead. We'll go in the same order, starting with the average actor.

```{r, warning = F, message = F}
nd <-
  tibble(prosoc_left = c(0, 1, 0, 1),
         condition   = c(0, 0, 1, 1))

(
  f <-
  fitted(b12.4,
         newdata = nd,
         re_formula = NA,
         probs = c(.1, .9)) %>% 
  as_tibble() %>% 
  bind_cols(nd) %>% 
  mutate(condition = str_c(prosoc_left, "/", condition) %>% 
           factor(., levels = c("0/0", "1/0", "0/1", "1/1")))
  )
```

You should notice a few things. Since `b12.4` is a multilevel model, it had three predictors: `prosoc_left`, `condition`, and `actor`. However, our `nd` data only included the first two of those predictors. The reason `fitted()` permitted that was because we set `re_formula = NA`. When you do that, you tell `fitted()` to ignore group-level effects (i.e., focus only on the fixed effects). This was our `fitted()` version of ignoring the `r_` vectors returned by `posterior_samples()`. Here's the plot.

```{r, fig.width = 2.5, fig.height = 3.25}
p4 <-
  f %>%
  ggplot(aes(x = condition, y = Estimate, group = 1)) +
  geom_ribbon(aes(ymin = Q10, ymax = Q90), fill = "blue") +
  geom_line(color = "orange1") +
  ggtitle("Average actor") +
  coord_cartesian(ylim = 0:1) +
  theme_fivethirtyeight() +
  theme(plot.title = element_text(size = 14, hjust = .5))

p4
```

For marginal of actor, we can continue using the same `nd` data. This time we’ll be sticking with the default `re_formula` setting, which will accommodate the multilevel nature of the model. However, we’ll also be adding `allow_new_levels = T` and `sample_new_levels = "gaussian"`. The former will allow us to marginalize across the specific actors in our data and the latter will instruct `fitted()` to use the multivariate normal distribution implied by the random effects. It’ll make more sense why I say *multivariate* normal by the end of the next chapter. For now, just go with it.

```{r, warning = F, message = F}
(
  f <-
  fitted(b12.4,
         newdata = nd,
         probs = c(.1, .9),
         allow_new_levels = T,
         sample_new_levels = "gaussian") %>% 
  as_tibble() %>% 
  bind_cols(nd) %>% 
  mutate(condition = str_c(prosoc_left, "/", condition) %>% 
           factor(., levels = c("0/0", "1/0", "0/1", "1/1")))
  )
```

Here's our `fitted()`-based marginal of `actor` plot.

```{r, fig.width = 2.5, fig.height = 3.25}
p5 <-
  f %>%
  ggplot(aes(x = condition, y = Estimate, group = 1)) +
  geom_ribbon(aes(ymin = Q10, ymax = Q90), fill = "blue") +
  geom_line(color = "orange1") +
  ggtitle("Marginal of actor") +
  coord_cartesian(ylim = 0:1) +
  theme_fivethirtyeight() +
  theme(plot.title = element_text(size = 14, hjust = .5))

p5
```

For the simulated actors plot, we’ll just amend our process from the last one. This time we’re setting `summary = F`, in order to keep the iteration-specific results, and setting `nsamples = n_sim`. `n_sim` is just a name for the number of actors we’d like to simulate (i.e., 50, as in the text).

```{r, warning = F, message = F}
# how many simulated actors would you like?
n_sim <- 50

(
  f <-
  fitted(b12.4,
         newdata = nd,
         probs = c(.1, .9),
         allow_new_levels = T,
         sample_new_levels = "gaussian",
         summary = F,
         nsamples = n_sim) %>% 
  as_tibble() %>% 
  mutate(iter = 1:n_sim) %>% 
  gather(key, value, -iter) %>% 
  bind_cols(nd %>% 
              transmute(condition = str_c(prosoc_left, "/", condition) %>% 
                          factor(., levels = c("0/0", "1/0", "0/1", "1/1"))) %>% 
              expand(condition, iter = 1:n_sim))
  )
```

```{r, fig.width = 2.5, fig.height = 3.25}
p6 <-
  f %>%
  
  ggplot(aes(x = condition, y = value, group = iter)) +
  geom_line(alpha = 1/2, color = "blue") +
  ggtitle("50 simulated actors") +
  coord_cartesian(ylim = 0:1) +
  theme_fivethirtyeight() +
  theme(plot.title = element_text(size = 14, hjust = .5))

p6
```

Here they are altogether.

```{r, fig.width = 7.5, fig.height = 3.25}
grid.arrange(p4, p5, p6, ncol = 3)
```

### Focus and multilevel prediction.

First, let's load the `Kline` data.

```{r, message = F}
# prep data
library(rethinking)
data(Kline)
d <- Kline
```

Switch out the packages, once again.

```{r, message = F}
detach(package:rethinking, unload = T)
library(brms)
rm(Kline)
```

The statistical formula for our multilevel count model is

\begin{align*}
\text{total_tools}_i    & \sim \text{Poisson} (\mu_i) \\
\text{log} (\mu_i)      & = \alpha + \alpha_{\text{culture}_i} + \beta \text{log} (\text{population}_i) \\
\alpha                  & \sim \text{Normal} (0, 10) \\
\beta                   & \sim \text{Normal} (0, 1) \\
\alpha_{\text{culture}} & \sim \text{Normal} (0, \sigma_{\text{culture}}) \\
\sigma_{\text{culture}} & \sim \text{HalfCauchy} (0, 1) \\
\end{align*}

With brms, we don't actually need to make the `logpop` or `society` variables. We're ready to fit the multilevel `Kline` model with the data in hand.

```{r b12.6, cache = T, message = F, warning = F}
b12.6 <- 
  brm(data = d, family = poisson,
      total_tools ~ 0 + intercept + log(population) + 
        (1 | culture),
      prior = c(prior(normal(0, 10), class = b, coef = intercept),
                prior(normal(0, 1), class = b),
                prior(cauchy(0, 1), class = sd)),
      iter = 4000, warmup = 1000, cores = 3, chains = 3,
      seed = 12)
```

Note how we used the special `0 + intercept` syntax rather than using the default Intercept. This is because our predictor variable was not mean centered. For more info, see [here](https://github.com/paul-buerkner/brms/issues/114). Though we used the `0 + intercept` syntax for the fixed effect, it was not necessary for the random effect. Both ways work.

Here is the data-processing work for our variant of Figure 12.6.

```{r}
nd <- 
  tibble(population = seq(from = 1000, to = 400000, by = 5000),
         # to "simulate counterfactual societies, using the hyper-parameters" (p. 383), 
         # we'll plug a new island into the `culture` variable
         culture    = "my_island") 

p <-
  predict(b12.6,
          # this allows us to simulate values for our counterfactual island, "my_island"
          allow_new_levels = T,
          # here we explicitly tell brms we want to include the group-level effects
          re_formula = ~ (1 | culture),
          # from the brms manual, this uses the "(multivariate) normal distribution implied by 
          # the group-level standard deviations and correlations", which appears to be 
          # what McElreath did in the text.
          sample_new_levels = "gaussian",
          newdata = nd,
          probs = c(.015, .055, .165, .835, .945, .985)) %>%
  as_tibble() %>%
  bind_cols(nd)

p %>%  
  glimpse()
```

For a detailed discussion on this way of using `brms::predict()`, see [Andrew MacDonald’s great blogpost on this very figure](http://thestudyofthehousehold.com/2018/02/13/2018-02-13-easily-made-fitted-and-predicted-values-made-easy/). Here’s what we’ve been working for:

```{r, fig.width = 3.75, fig.height = 3.5}
p %>%
  ggplot(aes(x = log(population), y = Estimate)) +
  geom_ribbon(aes(ymin = Q1.5,  ymax = Q98.5), fill = "orange2", alpha = 1/3) +
  geom_ribbon(aes(ymin = Q5.5,  ymax = Q94.5), fill = "orange2", alpha = 1/3) +
  geom_ribbon(aes(ymin = Q16.5, ymax = Q83.5), fill = "orange2", alpha = 1/3) +
  geom_line(color = "orange4") +
  geom_text(data = d, aes(y = total_tools, label = culture), 
            size = 2.33, color = "blue") +
  ggtitle("Total tools as a function of log(population)") +
  coord_cartesian(ylim = range(d$total_tools)) +
  theme_fivethirtyeight() +
  theme(plot.title = element_text(size = 12, hjust = .5))
```

Glorious.

> The envelope of predictions is a lot wider here than it was back in Chapter 10. This is a consequene of the varying intercepts, combined with the fact that there is much more variation in the data than a pure-Poisson model anticipates. (p. 384)

## ~~Summary~~ Bonus: `tidybayes::spread_draws()`

A big part of this chapter, both what McElreath focused on in the text and even our plotting digression a bit above, focused on how to combine the fixed effects of a multilevel with the group-level. Given some binomial variable, $\text{criterion}$, and some group term, $\text{grouping variable}$, we’ve learned the simple multilevel model follows a form like

\begin{align*}
\text{criterion}_i                & \sim \text{Binomial} (n_i \geq 1, p_i) \\
\text{logit} (p_i)                & = \alpha + \alpha_{\text{grouping variable}_i}\\
\alpha                            & \sim \text{Normal} (0, 1) \\
\alpha_{\text{grouping variable}} & \sim \text{Normal} (0, \sigma_{\text{grouping variable}}) \\
\sigma_{\text{grouping variable}} & \sim \text{HalfCauchy} (0, 1)
\end{align*}

and we’ve been grappling with the relation between the grand mean $\alpha$ and the group-level deviations $\alpha_{\text{grouping variable}}$. For situations where we have the `brms::brm()` model fit in hand, we’ve been playing with various ways to use the iterations, particularly with either the `posterior_samples()` method and the `fitted()/predict()` method. Both are great. But (a) we have other options, which I'd like to share, and (b) if you're like me, you probably need more practice than following along with the examples in the text. In this bonus section, we are going to introduce two simplified models and then practice working with combining the grand mean various combinations of the random effects.

For our first step, we'll introduce the models.

### Intercepts-only models with one or two grouping variables

If you recall, `b12.4` was our first multilevel model with the chimps data. We can retrieve the model formula like so.

```{r}
b12.4$formula
```
 
In addition to the model intercept and random effects for the individual chimps (i.e., `actor`), we also included fixed effects for the study conditions. For our bonus section, it'll be easier if we reduce this to a simple intercepts-only model with the sole `actor` grouping factor. That model will follow the form 

\begin{align*}
\text{pulled_left}_i  & \sim \text{Binomial} (n_i = 1, p_i) \\
\text{logit} (p_i)    & = \alpha + \alpha_{\text{actor}_i}\\
\alpha                & \sim \text{Normal} (0, 10) \\
\alpha_{\text{actor}} & \sim \text{Normal} (0, \sigma_{\text{actor}}) \\
\sigma_{\text{actor}} & \sim \text{HalfCauchy} (0, 1)
\end{align*}

Before we fit the model, you might recall that (a) we’ve already removed the `chimpanzees` data after saving the data as `d` and (b) we subsequently reassigned the `Kline` data to `d`. Instead of reloading the rethinking package to retrieve the `chimpanzees` data, we might also acknowledge that the data has also been saved within our `b12.4` fit object. [It’s easy to forget such things.]

```{r}
b12.4$data %>% 
  glimpse()
```

So there's no need to reload anything. Everything we need is already at hand. Let's fit the intercepts-only model.

```{r b12.7, cache = T, message = F, warning = F}
b12.7 <- 
  brm(data = b12.4$data, family = binomial,
      pulled_left | trials(1) ~ 1 + (1 | actor),
      prior = c(prior(normal(0, 10), class = Intercept),
                prior(cauchy(0, 1), class = sd)),
      iter = 5000, warmup = 1000, chains = 4, cores = 4,
      control = list(adapt_delta = 0.95),
      seed = 12)
```

Here's the model summary:

```{r}
print(b12.7)
```

Now recall that our competing cross-classified model, `b12.5` added random effects for the trial blocks. Here was that formula.

```{r}
b12.5$formula
```

And, of course, we can retrieve the data from that model, too.

```{r}
b12.5$data %>% 
  glimpse()
```

It’s the same data we used from the `b12.4` model, but with the addition of the `block` index. With those data in hand, we can fit the intercepts-only version of our cross-classified model. This model formula follows the form

\begin{align*}
\text{pulled_left}_i  & \sim \text{Binomial} (n_i = 1, p_i) \\
\text{logit} (p_i)    & = \alpha + \alpha_{\text{actor}_i} + \alpha_{\text{block}_i}\\
\alpha                & \sim \text{Normal} (0, 10) \\
\alpha_{\text{actor}} & \sim \text{Normal} (0, \sigma_{\text{actor}}) \\
\alpha_{\text{block}} & \sim \text{Normal} (0, \sigma_{\text{block}}) \\
\sigma_{\text{actor}} & \sim \text{HalfCauchy} (0, 1) \\
\sigma_{\text{block}} & \sim \text{HalfCauchy} (0, 1)
\end{align*}

Fit the model.

```{r b12.8, cache = T, message = F, warning = F}
b12.8 <- 
  brm(data = b12.5$data, family = binomial,
      pulled_left | trials(1) ~ 1 + (1 | actor) + (1 | block),
      prior = c(prior(normal(0, 10), class = Intercept),
                prior(cauchy(0, 1), class = sd)),
      iter = 5000, warmup = 1000, chains = 4, cores = 4,
      control = list(adapt_delta = 0.95),
      seed = 12)
```

Here's the summary.

```{r}
print(b12.8)
```

Now we’ve fit our two intercepts-only models, let’s get to the heart of this section. We are going to practice four methods for working with the posterior samples. Each method will revolve around a different primary function. In order, they are

* `brms::posterior_samples()`
* `brms::coef()`
* `brms::fitted()`
* `tidybayes::spread_draws()`

We’ve already had some practice with the first three, but I hope this section will make them even more clear. The `tidybayes::spread_draws()` method will be new, to us. I think you'll find it’s a handy alternative.

With each of the four methods, we’ll practice three different model summaries.

* Getting the posterior draws for the `actor`-level estimates from the `b12.7` model
* Getting the posterior draws for the `actor`-level estimates from the cross-classified `b12.8` model, averaging over the levels of `block`
* Getting the posterior draws for the `actor`-level estimates from the cross-classified `b12.8` model, based on `block == 1`

So to be clear, our goal is to accomplish those three tasks with four methods, each of which should yield equivalent results.

### `brms::posterior_samples()`

To warm up, let’s take a look at the structure of the `posterior_samples()` output for the simple `b12.7` model.

```{r}
posterior_samples(b12.7) %>% str()
```

The `b_Intercept` vector corresponds to the $\alpha$ term in the statistical model. The second vector, `sd_actor__Intercept`, corresponds to the $\sigma_{\text{actor}}$ term. And the next 7 vectors beginning with the `r_actor` suffix are the $\alpha_{\text{actor}}$ deviations from the grand mean, $\alpha$. Thus if we wanted to get the model-implied probability for our first chimp, we'd add `b_Intercept` to `r_actor[1,Intercept]` and then take the inverse logit.

```{r}
posterior_samples(b12.7) %>% 
  transmute(`chimp 1's average probability of pulling left` = (b_Intercept + `r_actor[1,Intercept]`) %>% inv_logit_scaled()) %>% 
  head()
```

To complete our first task, then, of getting the posterior draws for the `actor`-level estimates from the `b12.7` model, we can do that in bulk.

```{r}
p1 <-
  posterior_samples(b12.7) %>% 
  transmute(`chimp 1's average probability of pulling left` = b_Intercept + `r_actor[1,Intercept]`,
            `chimp 2's average probability of pulling left` = b_Intercept + `r_actor[2,Intercept]`,
            `chimp 3's average probability of pulling left` = b_Intercept + `r_actor[3,Intercept]`,
            `chimp 4's average probability of pulling left` = b_Intercept + `r_actor[4,Intercept]`,
            `chimp 5's average probability of pulling left` = b_Intercept + `r_actor[5,Intercept]`,
            `chimp 6's average probability of pulling left` = b_Intercept + `r_actor[6,Intercept]`,
            `chimp 7's average probability of pulling left` = b_Intercept + `r_actor[7,Intercept]`) %>% 
  mutate_all(inv_logit_scaled)

str(p1)
```

One of the things I really like about this method is the `b_Intercept + r_actor[i,Intercept]` part of the code makes it very clear, to me, how the `porterior_samples()` columns correspond to the statistical model, $\text{logit} (p_i) = \alpha + \alpha_{\text{actor}_i}$. This method easily extends to our next task, getting the posterior draws for the `actor`-level estimates from the cross-classified `b12.8` model, averaging over the levels of `block`. In fact, other than switching out `b12.7` for `b12.8`, the method is identical.

```{r}
p2 <-
  posterior_samples(b12.8) %>% 
  transmute(`chimp 1's average probability of pulling left` = b_Intercept + `r_actor[1,Intercept]`,
            `chimp 2's average probability of pulling left` = b_Intercept + `r_actor[2,Intercept]`,
            `chimp 3's average probability of pulling left` = b_Intercept + `r_actor[3,Intercept]`,
            `chimp 4's average probability of pulling left` = b_Intercept + `r_actor[4,Intercept]`,
            `chimp 5's average probability of pulling left` = b_Intercept + `r_actor[5,Intercept]`,
            `chimp 6's average probability of pulling left` = b_Intercept + `r_actor[6,Intercept]`,
            `chimp 7's average probability of pulling left` = b_Intercept + `r_actor[7,Intercept]`) %>% 
  mutate_all(inv_logit_scaled)

str(p2)
```

The reason we can still get away with this is because the grand mean in the `b12.8` model is the grand mean across all levels of `actor` and `block`. AND it’s the case that the `r_actor` and `r_block` vectors returned by `posterior_samples(b12.8)` are all in deviation metrics--execute `posterior_samples(b12.8) %>% glimpse()` if it will help you follow along. So if we simply leave out the `r_block` vectors, we are ignoring the specific `block`-level deviations, effectively averaging over them.

Now for our third task, we’ve decided we wanted to retrieve the posterior draws for the `actor`-level estimates from the cross-classified `b12.8` model, based on `block == 1`. To get the chimp-specific estimates for the first `block`, we simply add `+ r_block[1,Intercept]` to the end of each formula.

```{r}
p3 <-
  posterior_samples(b12.8) %>% 
  transmute(`chimp 1's average probability of pulling left` = b_Intercept + `r_actor[1,Intercept]` + `r_block[1,Intercept]`,
            `chimp 2's average probability of pulling left` = b_Intercept + `r_actor[2,Intercept]` + `r_block[1,Intercept]`,
            `chimp 3's average probability of pulling left` = b_Intercept + `r_actor[3,Intercept]` + `r_block[1,Intercept]`,
            `chimp 4's average probability of pulling left` = b_Intercept + `r_actor[4,Intercept]` + `r_block[1,Intercept]`,
            `chimp 5's average probability of pulling left` = b_Intercept + `r_actor[5,Intercept]` + `r_block[1,Intercept]`,
            `chimp 6's average probability of pulling left` = b_Intercept + `r_actor[6,Intercept]` + `r_block[1,Intercept]`,
            `chimp 7's average probability of pulling left` = b_Intercept + `r_actor[7,Intercept]` + `r_block[1,Intercept]`) %>% 
  mutate_all(inv_logit_scaled)

str(p3)
```

Again, I like this method because of how close the wrangling code within `transmute()` is to the statistical model formula. I wrote a lot of code like this in my early days of working with these kinds of models, and I think the pedagogical insights were helpful. But this method has its limitations. It works fine if you're working with some small number of groups. But that's a lot of repetitious code and it would be utterly un-scalable to situations where you have 50 or 500 levels in your grouping variable. We need alternatives. 

### `brms::coef()`

First, let’s review what the `coef()` function returns. 

```{r}
coef(b12.7)
```

By default, we get the familiar summaries for mean performances for each of our seven chimps. These, of course, are in the log-odds metric and simply tacking on `inv_logit_scaled()` isn’t going to fully get the job done. So to get things in the probability metric, we’ll want to first set `summary = F` in order to work directly with un-summarized samples and then wrangle quite a bit. Part of the wrangling challenge is because `coef()` returns a list, rather than a data frame. With that in mind, the code for our first task of getting the posterior draws for the `actor`-level estimates from the `b12.7` model looks like so.

```{r}
c1 <-
  coef(b12.7, summary = F)$actor[, , ] %>%
  as_tibble() %>% 
  gather() %>% 
  mutate(key   = str_c("chimp ", key, "'s average probability of pulling left"),
         value = inv_logit_scaled(value),
         # we need an iteration index for `spread()` to work properly
         iter  = rep(1:16000, times = 7)) %>% 
  spread(key = key, value = value)

str(c1)
```

So with this method, you get a little practice with three-dimensional indexing, which is a good skill to have. Now let’s extend it to our second task, getting the posterior draws for the `actor`-level estimates from the cross-classified `b12.8` model, averaging over the levels of `block`.

```{r}
c2 <-
  coef(b12.8, summary = F)$actor[, , ] %>%
  as_tibble() %>% 
  gather() %>% 
  mutate(key   = str_c("chimp ", key, "'s average probability of pulling left"),
         value = inv_logit_scaled(value),
         iter  = rep(1:16000, times = 7)) %>% 
  spread(key = key, value = value)

str(c2)
```

As with our `posterior_samples()` method, this code was near identical to the block, above. All we did was switch out `b12.7` for `b12.8`. [Okay, we removed a line of annotations. But that doesn't really count.] We should point something out, though. Consider what `coef()` yields when working with a cross-classified model.

```{r}
coef(b12.8)
```

Now we have a list of two elements, one for `actor` and one for `block`. What might not be immediately obvious is that the summaries returned by one grouping level are based off of averaging over the other. Although this made our second task easy, it provides a challenge for our third task, getting the posterior draws for the `actor`-level estimates from the cross-classified `b12.8` model, based on `block == 1`. To accomplish that, we’ll need to bring in `ranef()`. Let’s review what that returns.

```{r}
ranef(b12.8)
```

The format of the `ranef()` output is identical to that from `coef()`. However, the summaries are in the deviance metric. They’re all centered around zero, which corresponds to the part of the statistical model that specifies how $\alpha_{\text{block}} \sim \text{Normal} (0, \sigma_{\text{block}})$. So then, if we want to continue using our `coef()` method, we’ll need to augment it with `ranef()` to accomplish our last task.

```{r}
c3 <-
  coef(b12.8, summary = F)$actor[, , ] %>%
  as_tibble() %>% 
  gather() %>% 
  # here we add in the `block == 1` deviations from the grand mean
  mutate(value = value + ranef(b12.8, summary = F)$block[, 1, ] %>% rep(., times = 7)) %>% 
  mutate(key   = str_c("chimp ", key, "'s average probability of pulling left"),
         value = inv_logit_scaled(value),
         iter  = rep(1:16000, times = 7)) %>% 
  spread(key = key, value = value)

str(c3)
```

One of the nicest things about the `coef()` method is how is scales well. This code is no more burdensome for 5 group levels than it is for 5000. It’s also a post-processing version of the distinction McElreath made on page 372 between the two equivalent ways you might define a Gaussian:

$$\text{Normal}(10, 1)$$
and

$$10 + \text{Normal}(0, 1)$$

Conversely, it can be a little abstract. Let’s keep expanding our options. 

### `brms::fitted()`

As is often the case, we're going to want to define our predictor values for `fitted()`.

```{r}
(nd <- b12.7$data %>% distinct(actor))
```

Now we have our new data, `nd`, here's how we might use `fitted()` to accomplish our first task, getting the posterior draws for the `actor`-level estimates from the `b12.7` model.

```{r, warning = F, message = F}
f1 <-
  fitted(b12.7,
         newdata = nd,
         summary = F,
         # within `fitted()`, this line does the same work that
         # `inv_logit_scaled()` did with the other two methods
         scale = "response") %>% 
  as_tibble() %>% 
  set_names(str_c("chimp ", 1:7, "'s average probability of pulling left"))

str(f1)
```

This scales reasonably well. But might not work well if the vectors you wanted to rename didn’t follow a serial order, like ours. If you’re willing to pay with a few more lines of wrangling code, this method is more general, but still scalable.

```{r, warning = F, message = F}
f1 <-
  fitted(b12.7,
         newdata = nd,
         summary = F,
         scale = "response") %>% 
  as_tibble() %>% 
  # you'll need this line to make the `spread()` line work properly
  mutate(iter = 1:n()) %>% 
  gather(key, value, -iter) %>% 
  mutate(key = str_replace(key, "V", "chimp ")) %>% 
  mutate(key = str_c(key, "'s average probability of pulling left")) %>% 
  spread(key = key, value = value)

str(f1)
```

Now unlike with the previous two methods, our `fitted()` method will not allow us to simply switch out `b12.7` for `b12.8` to accomplish our second task of getting the posterior draws for the `actor`-level estimates from the cross-classified `b12.8` model, averaging over the levels of `block`. This is because when we use `fitted()` in combination with its `newdata` argument, the function expects us to define values for all the predictor variables in the formula. Because the `b12.8` model has both `actor` and `block` grouping variables as predictors, the default requires we include both in our new data. But if we were to specify a value for `block` in the `nd` data, we would no longer be averaging over the levels of `block` anymore; we’d be selecting one of the levels of `block` in particular, which we don’t yet want to do. Happily, `brms::fitted()` has a `re_formula` argument. If we would like to average out `block`, we simply drop it from the formula. Here’s how to do so.

```{r, warning = F, message = F}
f2 <-
  fitted(b12.8,
         newdata = nd,
         # this line allows us to average over the levels of `block`
         re_formula = pulled_left ~ 1 + (1 | actor),
         summary = F,
         scale = "response") %>% 
  as_tibble() %>% 
  set_names(str_c("chimp ", 1:7, "'s average probability of pulling left"))

str(f2)
```

If we want to use `fitted()` for our third task of getting the posterior draws for the `actor`-level estimates from the cross-classified `b12.8` model, based on `block == 1`,  we’ll need to augment our `nd` data.

```{r}
(
  nd <-
  b12.8$data %>% 
  distinct(actor) %>% 
  mutate(block = 1)
  )
```

This time, we no longer need that `re_formula` argument.

```{r, warning = F, message = F}
f3 <-
  fitted(b12.8,
         newdata = nd,
         summary = F,
         scale = "response") %>% 
  as_tibble() %>% 
  set_names(str_c("chimp ", 1:7, "'s average probability of pulling left"))

str(f3)
```

Let's learn one more option.

### `tidybayes::spread_draws()`

Up till this point, we’ve really only used the tidybayes package for plotting (e.g., with `geom_halfeyeh()`) and summarizing (e.g., with `median_qi()`). But tidybayes is more general; it offers a handful of convenience functions for wrangling posterior draws from a tidyverse perspective. One such function is `spread_draws()`, which you can learn all about in Matthew Kay's vignette [*Extracting and visualizing tidy draws from brms models*](https://mjskay.github.io/tidybayes/articles/tidy-brms.html). Let's take a look at how we'll be using it.

```{r, warning = F, message = F}
library(tidybayes)

b12.7 %>%
  spread_draws(b_Intercept, r_actor[actor,])
```

First, notice `tidybayes::spread_draws()` took the model fit itself, `b12.7`, as input. No need for `posterior_samples()`. Now, notice we fed it two additional arguments. By the first argument, we that requested `spead_draws()` extract the posterior samples for the `b_Intercept`. By the second argument, `r_actor[actor,]`, we instructed `spead_draws()` to extract all the random effects for the `actor` variable. Also notice how within the brackets `[]` we specified `actor`, which then became the name of the column in the output that indexed the levels of the grouping variable `actor`. By default, the code returns the posterior samples for all the levels of `actor`. However, had we only wanted those from chimps #1 and #3, we might use typical tidyverse-style indexing. E.g.,

```{r}
b12.7 %>%
  spread_draws(b_Intercept, r_actor[actor,]) %>% 
  filter(actor %in% c(1, 3))
```

Also notice those first three columns. By default, `spread_draws()` extracted information about which Markov chain a given draw was from, which iteration a given draw was within a given chain, and which draw from an overall standpoint. If it helps to keep track of which vector indexed what, consider this.

```{r}
b12.7 %>%
  spread_draws(b_Intercept, r_actor[actor,]) %>% 
  ungroup() %>% 
  select(.chain:.draw) %>% 
  gather() %>% 
  group_by(key) %>%
  summarise(min = min(value),
            max = max(value))
```

Above we simply summarized each of the three variables by their minimum and maximum values. If you recall that we fit `b12.7` with four Markov chains, each with 4000 post-warmup iterations, hopefully it’ll make sense what each of those three variables index. 

Now we’ve done a little clarification, let’s use `spread_draws()` to accomplish our first task, getting the posterior draws for the `actor`-level estimates from the `b12.7` model.

```{r}
s1 <-
  b12.7 %>%
  spread_draws(b_Intercept, r_actor[actor,]) %>% 
  mutate(p = (b_Intercept + r_actor) %>% inv_logit_scaled()) %>% 
  select(.draw, actor, p) %>% 
  ungroup() %>% 
  mutate(actor = str_c("chimp ", actor, "'s average probability of pulling left")) %>% 
  spread(value = p, key = actor)

str(s1)
```

The method remains essentially the same for accomplishing our second task, getting the posterior draws for the `actor`-level estimates from the cross-classified `b12.8` model, averaging over the levels of `block`.

```{r}
s2 <-
  b12.8 %>%
  spread_draws(b_Intercept, r_actor[actor,]) %>% 
  mutate(p = (b_Intercept + r_actor) %>% inv_logit_scaled()) %>% 
  select(.draw, actor, p) %>% 
  ungroup() %>% 
  mutate(actor = str_c("chimp ", actor, "'s average probability of pulling left")) %>% 
  spread(value = p, key = actor)

str(s2)
```

To accomplish our third task, we augment the `spread_draws()` and first `mutate()` lines, and add a `filter()` line between them.

```{r}
s3 <-
  b12.8 %>%
  spread_draws(b_Intercept, r_actor[actor,], r_block[block,]) %>% 
  filter(block == 1) %>% 
  mutate(p = (b_Intercept + r_actor + r_block) %>% inv_logit_scaled()) %>% 
  select(.draw, actor, p) %>% 
  ungroup() %>% 
  mutate(actor = str_c("chimp ", actor, "'s average probability of pulling left")) %>% 
  spread(value = p, key = actor)

str(s3)
```

Hopefully working through these examples gave you some insight on the relation between fixed and random effects within multilevel models, and perhaps added to your posterior-iteration-wrangling toolkit. 

## Reference {-}

[McElreath, R. (2016). *Statistical rethinking: A Bayesian course with examples in R and Stan.* Chapman & Hall/CRC Press.](https://xcelab.net/rm/statistical-rethinking/)

## Session info {-}

```{r}
sessionInfo()
```

```{r, warning = F, echo = F}
rm(d, b12.1, b12.2, post, post_mdn, b12.2.e, a, sigma, n_ponds, dsim, b12.3, n_iter, p_partpool, dfline, b12.3_new, two_gaussians, b12.4, b12.5, nd, chimp, new_dsim, chimp_2_fitted, chimp_2_d, my_fitted, chimp_5_my_fitted, chimp_5_d, post_average_actor, ran_ef, fix_ef, ran_and_fix_ef, marginal_effects, p1, p2, p3, f, p4, p5, p6, p, w, n_sim, b12.6, b12.7, b12.8, c1, c2, c3, f1, f2, f3, s1, s2, s3)
```

```{r, echo = F, message = F, warning = F, results = "hide"}
pacman::p_unload(pacman::p_loaded(), character.only = TRUE)

ggplot2::theme_set(ggplot2::theme_grey())

bayesplot::color_scheme_set("blue")
```
