---
title: "Ch. 10 Counting and Classification"
author: "A Solomon Kurz"
date: "`r format(Sys.Date())`"
output:
  html_document:
    code_folding: show
    keep_md: TRUE
---

## 10.1. Binomial regression

The `chimpanzees` data:

```{r, message = F, warning = F}
library(rethinking)
data(chimpanzees)
d <- chimpanzees
```

Switching from rethinking to brms.

```{r, message = F, warning = F,}
detach(package:rethinking)
library(brms)
rm(chimpanzees)
```
 
We start with the simple intercept-only logistic regression model.

```{r, message = F, warning = F, results = "hide"}
b10.1 <-
  brm(data = d, family = binomial,
      pulled_left ~ 1,
      prior = c(set_prior("normal(0, 10)", class = "Intercept")))
```

```{r, message = F, warning = F}
library(tidyverse)
 
fixef(b10.1) %>%
  round(digits = 2)
```

On page 81, [Gelman and Hill (2007)](http://www.stat.columbia.edu/~gelman/arm/) give the formula for `invlogit()`, which is our alternative to the `logistic()` function in rethinking.

```{r}
invlogit <- function(x){1/(1 + exp(-x))}
```

It's easy to use.

```{r, message = F}
c(.18, .46) %>%
  invlogit()

fixef(b10.1) %>%
  invlogit()
```

The next two chimp models add predictors.

```{r, message = F, warning = F, results = "hide"}
b10.2 <-
  brm(data = d, family = binomial,
      pulled_left ~ 1 + prosoc_left,
      prior = c(set_prior("normal(0, 10)", class = "Intercept"),
                set_prior("normal(0, 10)", class = "b")))

b10.3 <-
  brm(data = d, family = binomial,
      pulled_left ~ 1 + prosoc_left + condition:prosoc_left ,
      prior = c(set_prior("normal(0, 10)", class = "Intercept"),
                set_prior("normal(0, 10)", class = "b")))
```

Before comparing our models, we'll first save their WAIC estimates as objects. These will come in handy in just a bit.

```{r, message = F}
w_b10.1 <- waic(b10.1)
w_b10.2 <- waic(b10.2)
w_b10.3 <- waic(b10.3)

compare_ic(w_b10.1, w_b10.2, w_b10.3)
```

For this manuscript, we'll take our color scheme from the [wesanderson package](https://cran.r-project.org/web/packages/wesanderson/index.html)'s *Moonrise2* palette.

```{r, message = F, fig.width = 3, fig.height = 1}
# install.packages("wesanderson", dependencies = T)
library(wesanderson)

wes_palette("Moonrise2")

wes_palette("Moonrise2")[1:4]
```

We'll also take a few formatting cues from [Edward Tufte](https://www.edwardtufte.com/tufte/books_vdqi), curtesy of the [ggthemes package](https://cran.r-project.org/web/packages/ggthemes/index.html). The `theme_tufte()` function will change the default font and remove some chart junk. The `theme_set()` function, below, will make these adjustments the default for all subsequent ggplot2 plots. To undo this, just code `theme_set(theme_default())`.

```{r, message = F, warning = F}
library(ggthemes)
library(bayesplot)

theme_set(theme_default() + 
            theme_tufte() +
            theme(plot.background = element_rect(fill = wes_palette("Moonrise2")[3],
                                                 color = wes_palette("Moonrise2")[3])))
```

Finally, here's our WAIC plot.

```{r, fig.width = 4.5, fig.height = 1.25}
tibble(model = c("b10.1", "b10.2", "b10.3"),
       waic  = c(w_b10.1$waic, w_b10.2$waic, w_b10.3$waic),
       se    = c(w_b10.1$se_waic, w_b10.2$se_waic, w_b10.3$se_waic)) %>%

  ggplot() +
  geom_pointrange(aes(x = reorder(model, -waic), y = waic,
                      ymin = waic - se,
                      ymax = waic + se,
                      color = model),
                  shape = 16) +
  scale_color_manual(values = wes_palette("Moonrise2")[c(1:2, 4)]) +
  coord_flip() +
  labs(x = NULL, y = NULL,
       title = "WAIC") +
  theme(axis.ticks.y = element_blank(),
        legend.position = "none")
```

The estimates from model `b10.3`, which might not have the lowest WAIC value, but is the one most clearly corresponding to the structure of the experiment:

```{r}
print(b10.3)
```

Here's what the odds are multiplied by:

```{r, message = F}
fixef(b10.3)[2] %>%
  exp()
```

Given an estimated value of 4, the probability of a pull, all else equal, would be:

```{r}
invlogit(4)
```

Adding the coefficient, `fixef(b10.3)[2]`, would yield:

```{r}
(4 + fixef(b10.3)[2]) %>%
  invlogit()
```

For our variant of Figure 10.2., we use `brms::pp_average()` in place of `rethinking::ensemble()`.

```{r, fig.height = 2.75, fig.width = 3, warning = F, message = F}
# The combined fitted() results of the three models weighted by their WAICs
pp_a <- 
  pp_average(b10.1, b10.2, b10.3,
             weights = "waic",
             method = "fitted") %>% 
  as_tibble() %>% 
  bind_cols(b10.3$data) %>% 
  distinct(Estimate, `2.5%ile`, `97.5%ile`, condition, prosoc_left) %>% 
  mutate(x_axis = str_c(prosoc_left, condition, sep = "/")) %>%
  mutate(x_axis = factor(x_axis, levels = c("0/0", "1/0", "0/1", "1/1"))) %>% 
  rename(pulled_left = Estimate)

# The empirically-based summaries
d_plot <-
  d %>%
  group_by(actor, condition, prosoc_left) %>%
  summarise(pulled_left = mean(pulled_left)) %>%
  mutate(x_axis = str_c(prosoc_left, condition, sep = "/")) %>%
  mutate(x_axis = factor(x_axis, levels = c("0/0", "1/0", "0/1", "1/1")))

# The plot
ggplot() +
  geom_ribbon(data = pp_a,
              aes(x = x_axis,
                  ymin = `2.5%ile`, 
                  ymax = `97.5%ile`,
                  group = 0),
              fill = wes_palette("Moonrise2")[2]) +
  geom_line(data = pp_a,
            aes(x = x_axis, 
                y = pulled_left,
                group = 0)) +
  geom_line(data = d_plot,
            aes(x = x_axis, y = pulled_left, group = actor),
            color = wes_palette("Moonrise2")[1], size = 1/3) +
  scale_x_discrete(expand = c(.03, .03)) +
  coord_cartesian(ylim = 0:1) +
  labs(x = "prosoc_left/condition",
       y = "proportion pulled left") +
  theme(axis.ticks.x = element_blank())
```

McElreath didn't show the actual pairs plot in the text. Here's ours using `mcmc_pairs()`.

```{r, message = F, warning = F, fig.height = 4, fig.width = 4.5}
# this helps us set our custom color scheme
color_scheme_set(c(wes_palette("Moonrise2")[3], 
                   wes_palette("Moonrise2")[1], 
                   wes_palette("Moonrise2")[2], 
                   wes_palette("Moonrise2")[2], 
                   wes_palette("Moonrise2")[1], 
                   wes_palette("Moonrise2")[1]))

# the actual plot
mcmc_pairs(x = posterior_samples(b10.3),
           pars = c("b_Intercept", "b_prosoc_left", "b_prosoc_left:condition"),
           off_diag_args = list(size = 1/10, alpha = 1/6),
           diag_fun = "dens")
```

As McElreath asserted, the posterior looks pretty multivariate Gaussian.

Enclosing the `actor` variable within `factor()` will produce the indexing we need to get `actor`-specific intercepts.

```{r, message = F, warning = F}
b10.4 <-
  brm(data = d, family = binomial,
      pulled_left ~ 0 + factor(actor) + prosoc_left + condition:prosoc_left ,
      prior = c(set_prior("normal(0, 10)", class = "b")),
      chains = 2, iter = 2500, warmup = 500, cores = 2,
      control = list(adapt_delta = 0.9))
```

Within the tidyverse, `distinct()` yields the information you'd otherwise get from `unique()`.

```{r}
d %>%
  distinct(actor)
```

The posterior summary:

```{r}
print(b10.4)
```

Here's what the `posterior_samples` for `b10.4` looks like:

```{r}
post <- posterior_samples(b10.4)
 
post %>%
  glimpse()
```

Our variant of Figure 10.3.

```{r, fig.height = 2.75, fig.width = 3}
post %>%
  ggplot(aes(x = b_factoractor2)) +
  geom_density(color = "transparent",
               fill = wes_palette("Moonrise2")[1]) +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(x = NULL,
       title = "Actor 2's large and uncertain intercept",
       subtitle = "Once your log-odds are above, like, 4, it's all\npretty much a probability of 1.")
```

Figure 10.4., the idiographic trajectories for four of our chimps.

```{r, fig.height = 4, fig.width = 6, warning = F, message = F}
d_plot_4 <-
  d_plot %>%
  filter(actor %in% c(3, 5:7)) %>%
  ungroup() %>% 
  mutate(actor = str_c("actor ", actor))

ftd <-
  fitted(b10.4) %>% 
  as_tibble() %>% 
  bind_cols(b10.4$data) %>% 
  filter(actor %in% c(3, 5:7)) %>% 
  distinct(Estimate, `2.5%ile`, `97.5%ile`, condition, prosoc_left, actor) %>% 
  select(actor, everything()) %>% 
  mutate(actor = str_c("actor ", actor)) %>% 
  mutate(x_axis = str_c(prosoc_left, condition, sep = "/")) %>%
  mutate(x_axis = factor(x_axis, levels = c("0/0", "1/0", "0/1", "1/1"))) %>% 
  rename(pulled_left = Estimate)

  ggplot(data = ftd,
         aes(x = x_axis, y = pulled_left, group = actor)) +
  geom_ribbon(aes(x = x_axis,
                  ymin = `2.5%ile`, 
                  ymax = `97.5%ile`),
              fill = wes_palette("Moonrise2")[2]) +
  geom_line(aes(x = x_axis, 
                y = pulled_left)) +
  geom_line(data = d_plot_4,
            color = wes_palette("Moonrise2")[1], size = 1.25) +
  scale_x_discrete(expand = c(.03, .03)) +
  coord_cartesian(ylim = 0:1) +
  labs(x = "prosoc_left/condition",
       y = "proportion pulled left") +
  theme(axis.ticks.x = element_blank(),
        # color came from: http://www.color-hex.com/color/ccc591
        panel.background = element_rect(fill = "#d1ca9c",
                                        color = "transparent")) +
  facet_wrap(~actor)
```

### 10.1.2. Aggregated binomial: Chimpanzees again, condensed.

With the tidyverse, we use `group_by()` and `summarise()` to achieve what McElreath did with `aggregate()`.

```{r}
d_aggregated <-
  d %>%
  select(-recipient, -block, -trial, -chose_prosoc) %>%
  group_by(actor, condition, prosoc_left) %>%
  summarise(x = sum(pulled_left))

d_aggregated %>%
  slice(1:8)
```

To fit an aggregated binomial model in brms, we use the `[criterion] | trials()` syntax where the value that goes in `trials()` is either a fixed number, as in this case, or an index variable.

```{r, message = F, warning = F, results = "hide"}
b10.5 <-
  brm(data = d_aggregated, family = binomial,
      x | trials(18) ~ 1 + prosoc_left + condition:prosoc_left ,
      prior = c(set_prior("normal(0, 10)", class = "Intercept"),
                set_prior("normal(0, 10)", class = "b")),
      iter = 2500, warmup = 500, cores = 2, chains = 2)
```

We might compare `b10.3` with `b10.5` like this.

```{r, message = F, warning = F}
fixef(b10.3) %>% round(digits = 2)
fixef(b10.5) %>% round(digits = 2)
```

Close within rounding error.

### 10.1.3. Aggregated binomial: Graduate school admissions.

The infamous `UCBadmit` data:

```{r, message = F, warning = F}
# detach(package:brms)
library(rethinking)
data(UCBadmit)
d <- UCBadmit
```

Switching from rethinking to brms.

```{r, message = F, warning = F}
detach(package:rethinking)
library(brms)
rm(UCBadmit)

d
```

Here's our newly-constructed predictor, `male`, and the models that do/do not put it to work.

```{r, message = F, warning = F, results = "hide"}
d <- 
  d %>%
  mutate(male = ifelse(applicant.gender == "male", 1, 0))

b10.6 <-
  brm(data = d, family = binomial,
      admit | trials(applications) ~ 1 + male ,
      prior = c(set_prior("normal(0, 10)", class = "Intercept"),
                set_prior("normal(0, 10)", class = "b")),
      iter = 2500, warmup = 500, cores = 2, chains = 2)

b10.7 <-
  brm(data = d, family = binomial,
      admit | trials(applications) ~ 1,
      prior = c(set_prior("normal(0, 10)", class = "Intercept")),
      iter = 2500, warmup = 500, cores = 2, chains = 2)
```

The WAIC comparison:

```{r, message = F, warning = F}
waic(b10.6, b10.7)
```

If you use the LOO to compare the models instead of the WAIC, you'll run into warnings that multiple observations have "a pareto_k > 0.7." You might take this as an indication that these observations are unduly influential on the model parameters, something like what you'd get with a Cook's $D_{i}$ analysis. This suggests the model isn't a great fit for the data. For more on the topic, see [this discussion on stackoverflow.com in which several members of the Stan team weigh in](https://stackoverflow.com/questions/39578834/linear-model-diagnostics-for-bayesian-models-using-rstan/39595436), [this paper](https://arxiv.org/abs/1507.04544), the [loo reference manual](https://cran.r-project.org/web/packages/loo/loo.pdf), and [this presentation by Vehtari himself](https://www.youtube.com/watch?v=FUROJM3u5HQ&feature=youtu.be&a=).

But that's a bit of a tangent from the central thrust of this section. Here's a look at `b10.6`, the univariable model:

```{r}
print(b10.6)
```

```{r, warning = F, message = F}
fixef(b10.6)[2] %>%
  exp() %>%
  round(digits = 2)
```

The difference in admission probabilities.

```{r}
post <- posterior_samples(b10.6)

post %>%
  mutate(p_admit_male   = invlogit(b_Intercept + b_male),
         p_admit_female = invlogit(b_Intercept),
         diff_admit     = p_admit_male - p_admit_female) %>%
  summarise(`2.5%`  = quantile(diff_admit, probs = .025),
            `50%`   = median(diff_admit),
            `97.5%` = quantile(diff_admit, probs = .975))
```

Here's our version of Figure 10.5.

```{r, fig.height = 2.5, fig.width = 4}
d <-
  d %>%
  mutate(case = factor(1:12))

p_10.6 <- 
  predict(b10.6) %>% 
  as_tibble() %>% 
  bind_cols(d)

d_text <-
  d %>%
  group_by(dept) %>%
  summarise(case = mean(as.numeric(case)),
            admit = mean(admit/applications) + .05)

ggplot(data = d, aes(x = case, y = admit/applications)) +
  geom_pointrange(data = p_10.6, 
                  aes(y = Estimate/applications,
                      ymin = `2.5%ile`/applications ,
                      ymax = `97.5%ile`/applications),
                  color = wes_palette("Moonrise2")[1],
                  shape = 1, alpha = 1/3) +
  geom_point(color = wes_palette("Moonrise2")[2]) +
  geom_line(aes(group = dept),
            color = wes_palette("Moonrise2")[2]) +
  geom_text(data = d_text,
            aes(y = admit, label = dept),
            color = wes_palette("Moonrise2")[2],
            family = "serif") +
  coord_cartesian(ylim = 0:1) +
  labs(y = "Proportion admitted",
       title = "Posterior validation check") +
  theme(axis.ticks.x = element_blank())
```

As alluded to in all that LOO/pareto_k talk, above, this is not a great fit.

We don't need to coerce an index. But here are the models.

```{r, message = F, warning = F, results = "hide"}
b10.8 <-
  brm(data = d, family = binomial,
      admit | trials(applications) ~ 0 + dept,
      prior = c(set_prior("normal(0, 10)", class = "b")),
      iter = 2500, warmup = 500, cores = 2, chains = 2)

b10.9 <-
  brm(data = d, family = binomial,
      admit | trials(applications) ~ 0 + dept + male ,
      prior = c(set_prior("normal(0, 10)", class = "b")),
      iter = 2500, warmup = 500, cores = 2, chains = 2)
```

Here we compare all four models by the WAIC.

```{r, warning = F, message = F}
waic(b10.6, b10.7, b10.8, b10.9)
```

Our multivariable model, `b10.9`, looks like this:

```{r}
print(b10.9)
```

Here's our version of Figure 10.6.

```{r, fig.height = 2.5, fig.width = 4}
predict(b10.9) %>% 
  as_tibble() %>% 
  bind_cols(d) %>% 

ggplot(aes(x = case, y = admit/applications)) +
  geom_pointrange(aes(y = Estimate/applications,
                      ymin = `2.5%ile`/applications ,
                      ymax = `97.5%ile`/applications),
                  color = wes_palette("Moonrise2")[1],
                  shape = 1, alpha = 1/3) +
  geom_point(color = wes_palette("Moonrise2")[2]) +
  geom_line(aes(group = dept),
            color = wes_palette("Moonrise2")[2]) +
  geom_text(data = d_text,
            aes(y = admit, label = dept),
            color = wes_palette("Moonrise2")[2],
            family = "serif") +
  coord_cartesian(ylim = 0:1) +
  labs(y = "Proportion admitted",
       title = "Posterior validation check") +
  theme(axis.ticks.x = element_blank())
```

### 10.1.4. Fitting binomial regressions with `glm()`.

We're not here to learn frequentist code, so we're going to skip most of this section. But model `b.good` is worth fitting. Here's the data.

```{r}
# outcome and predictor almost perfectly associated
y <- c(rep(0, 10), rep(1, 10))

x <- c(rep(-1, 9), rep(1, 11))
```

The `b.good` model:

```{r, message = F, warning = F, results = "hide"}
b.good <-
  brm(data = list(y = y, x = x), family = binomial,
      y ~ 1 + x,
      prior = c(set_prior("normal(0, 10)", class = "Intercept"),
                set_prior("normal(0, 10)", class = "b")))
```

Our model summary will differ a bit from the one in the text. This is because of the MAP/HMC contrast.

```{r}
print(b.good)
```

Here's the `pairs()` plot McElreath excluded from the text:

```{r, fig.width = 3.25, fig.height = 3}
pairs(b.good)
```

This plot deserves and extensive quote from McElreath. "Inspecting the pairs plot ~~(not shown)~~ demonstrates just how subtle even simple models can be, once we start working with GLMs. I don't say this to scare the reader. But it's true that even simple models can behave in complicated ways. How you fit the model is part of the model, and in principle no GLM is safe for MAP estimation."

## 10.2. Poisson regression

We'll simulate our sweet count data.

```{r}
set.seed(9968400) # making the results reproducible

y <- rbinom(1e5, 1000, 1/1000)

y %>%
  mean()

y %>%
  var()
```

```{r, message = F, warning = F}
# detach(package:brms)
library(rethinking)
data(Kline)
d <- Kline
```

Switching from rethinking to brms.

```{r, message = F, warning = F}
detach(package:rethinking)
library(brms)
rm(Kline)

d
```

Here are our new columns.

```{r}
d <-
  d %>%
  mutate(log_pop = log(population),
         contact_high = ifelse(contact == "high", 1, 0))
```

Our first Poisson model!

```{r, message = F, warning = F, results = "hide"}
b10.10 <-
  brm(data = d, family = poisson,
      total_tools ~ 1 + log_pop + contact_high + contact_high:log_pop,
      prior = c(set_prior("normal(0, 100)", class = "Intercept"),
                set_prior("normal(0, 1)", class = "b")),
      iter = 3000, warmup = 1000, chains = 4, cores = 4)
```

```{r}
print(b10.10)
```

Instead of looking at the correlation point estimates, we might plot:

```{r}
post <-
  posterior_samples(b10.10)

post[1:4] %>%
  rename(b_interaction = `b_log_pop:contact_high`) %>%
  cor() %>%
  round(digits = 2)
```

How plausible is it a high-contact island will have more tools than a low-contact island?

```{r}
post <-
  post %>%
  mutate(lambda_high = exp(b_Intercept + b_contact_high + (b_log_pop + `b_log_pop:contact_high`)*8),
         lambda_low = exp(b_Intercept + b_log_pop*8),
         diff = lambda_high - lambda_low) 

post %>%
  summarise(sum = sum(diff > 0)/length(diff))
```

Quite.

Here we are, Figure 10.8.a.

```{r, fig.width = 3, fig.height = 2.9}
post %>%
  ggplot(aes(x = diff)) +
  geom_density(color = "transparent",
               fill = wes_palette("Moonrise2")[1]) +
  geom_vline(xintercept = 0, linetype = 2,
             color = wes_palette("Moonrise2")[2]) +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(x = "lambda_high - lambda_low")
```

I'm not sure how to elegantly insert the marginal posterior estimates as dot-and-line plots into Figure 10.8.b., so we'll just throw in a couple rug plots, instead.

```{r, fig.width = 3, fig.height = 2.9}
post %>%
  ggplot(aes(x = b_contact_high, y = `b_log_pop:contact_high`)) +
  geom_point(color = wes_palette("Moonrise2")[1],
             size = 1/10, alpha = 1/10) +
  geom_rug(color = wes_palette("Moonrise2")[1],
           size = 1/15, alpha = 1/8)
```

Here we deconstruct model `b10.10`, bit by bit.

```{r, message = F, warning = F, results = "hide"}
# no interaction
b10.11 <-
  brm(data = d, family = poisson,
      total_tools ~ 1 + log_pop + contact_high,
      prior = c(set_prior("normal(0, 100)", class = "Intercept"),
                set_prior("normal(0, 1)", class = "b")),
      iter = 3000, warmup = 1000, chains = 4, cores = 4)

# no contact rate
b10.12 <-
  brm(data = d, family = poisson,
      total_tools ~ 1 + log_pop,
      prior = c(set_prior("normal(0, 100)", class = "Intercept"),
                set_prior("normal(0, 1)", class = "b")),
      iter = 3000, warmup = 1000, chains = 4, cores = 4)

# no log-population
b10.13 <-
  brm(data = d, family = poisson,
      total_tools ~ 1 + contact_high,
      prior = c(set_prior("normal(0, 100)", class = "Intercept"),
                set_prior("normal(0, 1)", class = "b")),
      iter = 3000, warmup = 1000, chains = 4, cores = 4)

# intercept only
b10.14 <-
  brm(data = d, family = poisson,
      total_tools ~ 1,
      prior = c(set_prior("normal(0, 100)", class = "Intercept")),
      iter = 3000, warmup = 1000, chains = 4, cores = 4)
```

```{r, message = F, warning = F}
w_b10.10 <- waic(b10.10)
w_b10.11 <- waic(b10.11)
w_b10.12 <- waic(b10.12)
w_b10.13 <- waic(b10.13)
w_b10.14 <- waic(b10.14)
 
compare_ic(w_b10.10, w_b10.11, w_b10.12, w_b10.13, w_b10.14)
```

Our handmade WAIC plot.

```{r, fig.width = 4.5, fig.height = 1.25}
tibble(model = c("b10.10", "b10.11", "b10.12", "b10.13", "b10.14"),
       waic  = c(w_b10.10$waic, w_b10.11$waic, w_b10.12$waic, w_b10.13$waic, w_b10.14$waic),
       se    = c(w_b10.10$se_waic, w_b10.11$se_waic, w_b10.12$se_waic, w_b10.13$se_waic, w_b10.14$se_waic)) %>%

  ggplot() +
  geom_pointrange(aes(x = reorder(model, -waic), y = waic,
                      ymin = waic - se,
                      ymax = waic + se,
                      color = model),
                  shape = 16) +
  scale_color_manual(values = wes_palette("Moonrise2")[c(1, 2, 1, 1, 1)]) +
  coord_flip() +
  labs(x = NULL, y = NULL,
       title = "WAIC") +
  theme(axis.ticks.y = element_blank(),
        legend.position = "none")
```

Our version of Figure 10.9. Recall, to do an "ensemble" posterior prediction in brms, one uses the `pp_average()` function.

```{r, fig.width = 3.5, fig.height = 3.5}
nd <-
  tibble(log_pop = rep(seq(from = 6.5, 
                           to = 13, 
                           length.out = 50),
                       times = 2),
         contact_high = rep(0:1, each = 50))

ppa_10.9 <- 
  pp_average(b10.10, b10.11, b10.12,
             weights = "waic",
             method = "fitted",
             newdata = nd) %>%
  as_tibble() %>%
  bind_cols(nd)

ppa_10.9 %>%
  ggplot(aes(x = log_pop,
             group = contact_high)) +
  geom_ribbon(aes(ymin = `2.5%ile`,
                  ymax = `97.5%ile`,
                  fill = contact_high),
              alpha = 1/4) +
  geom_line(aes(y = Estimate, color = contact_high)) +
  geom_text(data = d, 
             aes(y = total_tools,
                 label = total_tools,
                 color = contact_high),
             size = 3.5) +
  coord_cartesian(xlim = c(7.1, 12.4),
                  ylim = c(12, 70)) +
    labs(x = "log population",
         y = "total tools",
         subtitle = "Blue is the high contact rate and black is the low.") +
  theme(legend.position = "none",
        panel.border = element_blank())
```

### 10.2.2. MCMC islands.

We fit our analogue to `m10.10stan`, `b10.10`, some time ago. 

```{r}
print(b10.10)
```

Here's the x-centered version.

```{r, message = F, warning = F, results = "hide"}
d <-
  d %>%
  mutate(log_pop_c = log_pop - mean(log_pop))

b10.10.c <-
  brm(data = d, family = poisson,
      total_tools ~ 1 + log_pop_c + contact_high + contact_high:log_pop_c,
      prior = c(set_prior("normal(0, 10)", class = "Intercept"),
                set_prior("normal(0, 10)", class = "b")),
      iter = 3000, warmup = 1000, chains = 4, cores = 4)
```

The results of our centering:

```{r}
print(b10.10.c)
```

Figure 10.10.a. 

```{r, message = F, warning = F, fig.height = 3.5, fig.width = 3.75}
# this helps us set our custom color scheme
color_scheme_set(c(wes_palette("Moonrise2")[3], 
                   wes_palette("Moonrise2")[1], 
                   wes_palette("Moonrise2")[2], 
                   wes_palette("Moonrise2")[2], 
                   wes_palette("Moonrise2")[1], 
                   wes_palette("Moonrise2")[1]))

# the actual plot
mcmc_pairs(x = posterior_samples(b10.10),
           pars = c("b_Intercept", "b_log_pop", "b_contact_high", "b_log_pop:contact_high"),
           off_diag_args = list(size = 1/10, alpha = 1/10),
           diag_fun = "dens")
```

Figure 10.10.b. 

```{r, message = F, warning = F, fig.height = 3.5, fig.width = 3.75}
mcmc_pairs(x = posterior_samples(b10.10.c),
           pars = c("b_Intercept", "b_log_pop_c", "b_contact_high", "b_log_pop_c:contact_high"),
           off_diag_args = list(size = 1/10, alpha = 1/10),
           diag_fun = "dens")
```

If you really want the correlation point estimates, `lowerCor()` from the [psych package](https://cran.r-project.org/web/packages/psych/index.html) gives a nice way to get the lower triangle of the matrix.

```{r, warning = F, message = F}
library(psych)

lowerCor(posterior_samples(b10.10)[, 1:4])
lowerCor(posterior_samples(b10.10.c)[, 1:4])
```

### 10.2.3. Example: Exposure and the offset.

Here we simulate our data.

```{r}
set.seed(3838) # making it reproducible 

num_days <- 30
y <- rpois(num_days, 1.5)
```

```{r}
set.seed(3838) # making it reproducible 

num_weeks <- 4
y_new <- rpois(num_weeks, 0.5*7)
```

Let's make them tidy.

```{r}
d <- 
  tibble(y = c(y, y_new), 
         days = c(rep(1, 30), rep(7, 4)),
         monastery = c(rep(0, 30), rep(1, 4)))

d
```

Here we compute the offset and fit the model. With the brms package, you use the `offset()` syntax, in which you put a pre-processed variable like `log_days` or the log of a variable, such as `log(days)`.

```{r, message = F, warning = F, results = "hide"}
d <-
  d %>%
  mutate(log_days = log(days))

b10.15 <-
  brm(data = d, family = poisson,
      y ~ 1 + offset(log_days) + monastery,
      prior = c(set_prior("normal(0, 100)", class = "Intercept"),
                set_prior("normal(0, 1)", class = "b")),
      iter = 2500, warmup = 500, cores = 2, chains = 2)
```

The model summary:

```{r}
print(b10.15)
```

```{r}
posterior_samples(b10.15) %>%
  mutate(lambda_old = exp(b_Intercept),
         lambda_new  = exp(b_Intercept + b_monastery)) %>%
  gather(key, value, -(b_Intercept:lp__)) %>%
  mutate(key = factor(key, levels = c("lambda_old", "lambda_new"))) %>%
  group_by(key) %>%
  summarise(Mean = mean(value) %>% round(digits = 2),
            StdDev = sd(value) %>% round(digits = 2),
            LL = quantile(value, probs = .025) %>% round(digits = 2),
            UL = quantile(value, probs = .975) %>% round(digits = 2)) 
```

## 10.3. Other count regressions

### 10.3.1. Multinomial.

More simulation.

```{r, warning = F, message = F}
detach(package:brms)
library(rethinking)

# simulate career choices among 500 individuals
N <- 500             # number of individuals
income <- 1:3        # expected income of each career
score <- 0.5*income  # scores for each career, based on income

# next line converts scores to probabilities
p <- softmax(score[1], score[2], score[3])

# now simulate choice
# outcome career holds event type values, not counts
career <- rep(NA, N)  # empty vector of choices for each individual

set.seed(2078)
# sample chosen career for each individual
for(i in 1:N) career[i] <- sample(1:3, size = 1, prob = p)
```

Here's what the data look like.

```{r, fig.width = 3, fig.height = 2.25}
career %>%
  as_tibble() %>%
  ggplot(aes(x = value %>% as.factor())) +
  geom_bar(size = 0, fill = wes_palette("Moonrise2")[2])
```

Here's my naive attempt to fit the model in brms.

```{r, message = F, warning = F}
detach(package:rethinking)
library(brms)

b10.16 <-
  brm(data = list(career = career), 
      family = categorical(link = "logit"),
      career ~ 1,
      prior = c(set_prior("normal(0, 5)", class = "Intercept")),
      iter = 2500, warmup = 500, cores = 2, chains = 2)
```

This differs from McElreath's `m10.16`. Most obviously, this has two parameters. McElreath's `m10.16` only has one. If you have experience with these models and know how to reproduce McElreath's results in brms, hit me up.

```{r}
print(b10.16)
```

```{r, warning = F, message = F}
detach(package:brms)
library(rethinking)

N <- 100

set.seed(2078)
# simulate family incomes for each individual
family_income <- runif(N)

# assign a unique coefficient for each type of event
b <- (1:-1)
career <- rep(NA, N)  # empty vector of choices for each individual

for (i in 1:N) {
    score <- 0.5*(1:3) + b*family_income[i]
    p <- softmax(score[1], score[2], score[3])
    career[i] <- sample(1:3, size = 1, prob = p)
}
```

Here's the brms version of McElreath's `m10.17`.

```{r, message = F, warning = F}
detach(package:rethinking)
library(brms)

b10.17 <-
  brm(data = list(career = career,
                family_income = family_income), 
      family = categorical(link = "logit"),
      career ~ 1 + family_income,
       prior = c(set_prior("normal(0, 5)", class = "Intercept"),
                 set_prior("normal(0, 5)", class = "b")),
      iter = 2500, warmup = 500, cores = 2, chains = 2)
```

Happily, these results cohere with the rethinking model.

```{r}
print(b10.17)
```

McElreath describes the parameters as "on a scale that is very hard to interpret (p. 325)." Indeed.

#### 10.3.1.2.

Back to Berkeley

```{r, warning = F, message = F}
library(rethinking)

data(UCBadmit)
d <- UCBadmit
rm(UCBadmit)

detach(package:rethinking)
library(brms)
```

```{r, warning = F, message = F}
# binomial model of overall admission probability
b_binom <-
  brm(data = d, family = binomial,
      admit | trials(applications) ~ 1,
      prior = c(set_prior("normal(0, 100)", class = "Intercept")),
      iter = 2000, warmup = 1000, cores = 3, chains = 3)

# Poisson model of overall admission rate and rejection rate
d <-
  d %>%
  mutate(rej = reject) # 'reject' is a reserved word

b_pois <-
  brm(data = d, family = poisson,
      cbind(admit, rej) ~ 1,
      prior = c(set_prior("normal(0, 100)", class = "Intercept")),
      iter = 2000, warmup = 1000, cores = 3, chains = 3)
```

Note, the `cbind()` syntax made `b_pois` a multivariate Poisson model. Starting with version 2.0.0., [brms supports a variety of multivariate models](https://cran.r-project.org/web/packages/brms/vignettes/brms_multivariate.html). Anyway, here are the implications of `b_pois`.

```{r, fig.height = 3, fig.width = 3.25}
post <- posterior_samples(b_pois)

post %>%
  ggplot(aes(x = exp(b_admit_Intercept))) +
  geom_density(fill = wes_palette("Moonrise2")[2], size = 0) +
  geom_vline(xintercept = mean(d$admit), color = wes_palette("Moonrise2")[1]) +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(x = "# applications",
       title = "Mean acceptance # across departments:",
       subtitle = "The density is the posterior distribution. The line is the\nvalue in the data.")

post %>%
  ggplot(aes(x = exp(b_rej_Intercept))) +
  geom_density(fill = wes_palette("Moonrise2")[1], size = 0) +
  geom_vline(xintercept = mean(d$rej), color = wes_palette("Moonrise2")[2]) +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(x = "# applications",
       title = "Mean rejection # across departments:",
       subtitle = "The density is the posterior distribution. The line is the\nvalue in the data.")
```

The model summaries:

```{r}
print(b_binom)
print(b_pois)
```

Here's the posterior mean for the probability of admission, based on `b_binom`.

```{r}
fixef(b_binom) %>%
  invlogit()
```

Happily, we get the same value within simulation error from model `b_pois`.

```{r}
k <- 
  fixef(b_pois) %>%
  as.numeric()

exp(k[1])/(exp(k[1]) + exp(k[2]))
```

### 10.3.2. Geometric.

```{r}
# simulate
N <- 100
set.seed(1028)
x <- runif(N)

set.seed(1028)
y <- rgeom(N, prob = invlogit(-1 + 2*x))
```

In case you're curious, here are the data.

```{r, fig.width = 4, fig.height = 2.5}
list(y = y, x = x) %>%
  as_tibble() %>%
  ggplot(aes(x = x, y = y)) +
  geom_point(size = 3/5, alpha = 2/3)
```

Our geometric model:

```{r, warning = F, message = F}
b10.18 <-
  brm(data = list(y = y, x = x), 
      family = geometric(link = "log"),
      y ~ 0 + intercept + x,
      prior = c(set_prior("normal(0, 10)", class = "b", coef = "intercept"),
                set_prior("normal(0, 1)", class = "b")),
      chains = 2, iter = 2500, warmup = 500, cores = 2)
```

The results:

```{r}
print(b10.18, digits = 2)
```

It appears brms uses a [different parameterization for the exponential distribution](https://cran.r-project.org/web/packages/brms/vignettes/brms_families.html) than rethinking does. Even though the parameters brms yielded look different from those in the text, their predictions describe the data well. Here's the `marginal_effects()` plot:

```{r, fig.width = 4, fig.height = 2.5}
plot(marginal_effects(b10.18),
     points = T,
     point_args = c(size = 3/5, alpha = 2/3),
     line_args = c(color = wes_palette("Moonrise2")[1],
                   fill = wes_palette("Moonrise2")[1]))
```



```{r, echo = F}
rm(d, b10.1, invlogit, b10.2, b10.3, w_b10.1, w_b10.2, w_b10.3, post, d_plot, b10.4, d_plot_4, d_aggregated, b10.5, b10.6, b10.7, p_10.6, d_text, b10.8, b10.9, y, x, b.good, b10.10, b10.11, b10.12, b10.13, b10.14, w_b10.10, w_b10.11, w_b10.12, w_b10.13, w_b10.14, nd, ppa_10.9, b10.10.c, num_days, num_weeks, y_new, b10.15, N, income, score, p, career, i, b10.16, family_income, b, b10.17, k, b_binom, b_pois, b10.18)
```

Note. The analyses in this document were done with:

* R           3.4.4
* RStudio     1.1.442
* rmarkdown   1.9
* rstan       2.17.3
* rethinking  1.59
* brms        2.1.9
* tidyverse   1.2.1
* wesanderson 0.3.2
* ggthemes    3.4.0
* bayesplot   1.4.0
* psych       1.7.3.21

## Reference
McElreath, R. (2016). *Statistical rethinking: A Bayesian course with examples in R and Stan.* Chapman & Hall/CRC Press.