---
title: "Ch. 5 Multivariate Linear Models"
author: "A Solomon Kurz"
date: "`r format(Sys.Date())`"
output:
  html_document
---

```{r set-options, echo = F, cache = F}
options(width = 100)
```

Let's get the data from McElreath's [rethinking package](http://xcelab.net/rm/statistical-rethinking/).

```{r, message = F, warning = F}
library(rethinking)
library(tidyverse)

data(WaffleDivorce)
d <- WaffleDivorce
rm(WaffleDivorce)
```

The data look like so:

```{r}
glimpse(d)
```

## `b5.1`

Here we'll officially standardize the predictor, `MedianAgeMarriage`.

```{r}
d <-
  d %>%
  mutate(MedianAgeMarriage_s = (MedianAgeMarriage - mean(MedianAgeMarriage)) /
           sd(MedianAgeMarriage))
```

The likelihood for model `b5.1` is 

$$
\begin{eqnarray}
\text{Divorce}_i & \sim & \text{Normal}(\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta \text{MedianAgeMarriage_s}_i \\
\alpha & \sim & \text{Normal}(0, 10) \\
\beta & \sim & \text{Normal}(0, 1) \\
\sigma & \sim & \text{Uniform}(0, 10)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2}
set.seed(5.1)
d %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, Divorce, MedianAgeMarriage_s), index = letters[1:4]) %>% 
  mutate(alpha       = rnorm(n = n(), mean = 0, sd = 10),
         beta        = rnorm(n = n(), mean = 0, sd = 1),
         sigma       = runif(n = n(), 0, 10)) %>% 
  mutate(sim_Divorce = rnorm(n = n(),
                             mean = alpha + beta * MedianAgeMarriage_s,
                             sd   = sigma)) %>% 
  
  ggplot(aes(x = Divorce, y = sim_Divorce)) +
  geom_abline(color = "white") +
  geom_point(alpha = 1/2, size = 1/2) +
  theme(panel.grid = element_blank()) +
  facet_wrap(~index, scales = "free_y", ncol = 4)
```

We're going to be making a lot of plots of that same structure, just with different variables on the x and the y. To streamline our code, let's make a custom plotting function, `make_scatter()`.

```{r}
make_scatter <- function(data, mapping, ...){
  ggplot(data, mapping) +
  geom_abline(color = "white") +
  geom_point(alpha = 1/2, size = 1/2) +
  theme(panel.grid = element_blank()) +
  facet_wrap(~index, scales = "free_y", ncol = 4)
}
```

We'll put it to use in subsequent plots. To learn more about wrapping custom plots into custom functions, check out Chapter 12 of Wickham's [*ggplot2, Elegant graphics for data analysis*](https://github.com/hadley/ggplot2-book).

## `b5.2`

Here we'll officially standardize the predictor, `MedianAgeMarriage`.

```{r}
d <-
  d %>%
  mutate(Marriage_s = (Marriage - mean(Marriage)) / sd(Marriage))
```

The likelihood for model `b5.2` is 

$$
\begin{eqnarray}
\text{Divorce}_i & \sim & \text{Normal}(\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta \text{Marriage_s}_i \\
\alpha & \sim & \text{Normal}(0, 10) \\
\beta & \sim & \text{Normal}(0, 1) \\
\sigma & \sim & \text{Uniform}(0, 10)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2}
set.seed(5.2)
d %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, Divorce, Marriage_s), index = letters[1:4]) %>% 
  mutate(alpha       = rnorm(n = n(), mean = 0, sd = 10),
         beta        = rnorm(n = n(), mean = 0, sd = 1),
         sigma       = runif(n = n(), 0, 10)) %>% 
  mutate(sim_Divorce = rnorm(n = n(), 
                             mean = alpha + beta * Marriage_s, 
                             sd   = sigma)) %>% 
  
  make_scatter(aes(x = Divorce, y = sim_Divorce))
```

## `b5.3`

Here's the likelihood for `b5.3`, the first multivariable model:

$$
\begin{eqnarray}
\text{Divorce}_i & \sim & \text{Normal}(\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta_1 \text{Marriage_s}_i + \beta_2 \text{MedianAgeMarriage_s}_i \\
\alpha & \sim & \text{Normal}(0, 10) \\
\beta_1 & \sim & \text{Normal}(0, 1) \\
\beta_2 & \sim & \text{Normal}(0, 1) \\
\sigma & \sim & \text{Uniform}(0, 10)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2}
set.seed(5.3)
d %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, Divorce, Marriage_s, MedianAgeMarriage_s), index = letters[1:4]) %>% 
  mutate(alpha       = rnorm(n = n(), mean = 0, sd = 10),
         beta_1      = rnorm(n = n(), mean = 0, sd = 1),
         beta_2      = rnorm(n = n(), mean = 0, sd = 1),
         sigma       = runif(n = n(), 0, 10)) %>% 
  mutate(sim_Divorce = rnorm(n = n(),
                             mean = alpha + beta_1 * Marriage_s + beta_2 * MedianAgeMarriage_s,
                             sd   = sigma)) %>% 
  
  make_scatter(aes(x = Divorce, y = sim_Divorce))
```

## `b5.4`

The next model, `b5.4`, was for the first predictor residual plot. So the criterion has changed. Here's the likelihood:

$$
\begin{eqnarray}
\text{Marriage_s}_i & \sim & \text{Normal}(\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta \text{MedianAgeMarriage_s}_i \\
\alpha & \sim & \text{Normal}(0, 10) \\
\beta & \sim & \text{Normal}(0, 1) \\
\sigma & \sim & \text{Uniform}(0, 10)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2}
set.seed(5.4)
d %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, Marriage_s, MedianAgeMarriage_s), index = letters[1:4]) %>% 
  mutate(alpha          = rnorm(n = n(), mean = 0, sd = 10),
         beta           = rnorm(n = n(), mean = 0, sd = 1),
         sigma          = runif(n = n(), 0, 10)) %>% 
  mutate(sim_Marriage_s = rnorm(n = n(),
                                mean = alpha + beta * MedianAgeMarriage_s, 
                                sd   = sigma)) %>% 
  
  make_scatter(aes(x = Marriage_s, y = sim_Marriage_s))
```

## `b5.4b`

We do the reverse for model `b5.4b`.

$$
\begin{eqnarray}
\text{MedianAgeMarriage_s}_i & \sim & \text{Normal}(\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta \text{Marriage_s}_i \\
\alpha & \sim & \text{Normal}(0, 10) \\
\beta & \sim & \text{Normal}(0, 1) \\
\sigma & \sim & \text{Uniform}(0, 10)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2}
set.seed(5.4)
d %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, Marriage_s, MedianAgeMarriage_s), index = letters[1:4]) %>% 
  mutate(alpha                   = rnorm(n = n(), mean = 0, sd = 10),
         beta                    = rnorm(n = n(), mean = 0, sd = 1),
         sigma                   = runif(n = n(), 0, 10)) %>% 
  mutate(sim_MedianAgeMarriage_s = rnorm(n = n(), 
                                         mean = alpha + beta * Marriage_s,
                                         sd   = sigma)) %>% 
  
  make_scatter(aes(x = MedianAgeMarriage_s, y = sim_MedianAgeMarriage_s))
```

## `b5.5`

```{r, message = F}
data(milk)
d <- milk
rm(milk)
```

Here are the new `milk` data.

```{r}
glimpse(d)
```

The likelihood for `b5.5` is

$$
\begin{eqnarray}
\text{kcal.per.g}_i & \sim & \text{Normal}(\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta \text{neocortex.perc}_i \\
\alpha & \sim & \text{Normal}(0, 100) \\
\beta & \sim & \text{Normal}(0, 1) \\
\sigma & \sim & \text{HalfCauchy}(0, 1)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2, warning = F, message = F}
set.seed(5.5)
d %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, kcal.per.g, neocortex.perc), index = letters[1:4]) %>% 
  mutate(alpha          = rnorm(n = n(), mean = 0, sd = 10),
         beta           = rnorm(n = n(), mean = 0, sd = 1),
         sigma          = rcauchy(n = n(), location = 0, scale = 10) %>% abs()) %>% 
  mutate(sim_kcal.per.g = rnorm(n = n(),
                                mean = alpha + beta * neocortex.perc,
                                sd   = sigma)) %>% 
  
  make_scatter(aes(x = kcal.per.g, y = sim_kcal.per.g))
```

## `b5.6`

The next model requires a new variable.

```{r}
d <-
  d %>%
  mutate(log_mass = log(mass))
```

$$
\begin{eqnarray}
\text{kcal.per.g}_i & \sim & \text{Normal}(\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta \text{log_mass}_i \\
\alpha & \sim & \text{Normal}(0, 100) \\
\beta & \sim & \text{Normal}(0, 1) \\
\sigma & \sim & \text{Uniform}(0, 1)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2}
set.seed(5.6)
d %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, kcal.per.g, log_mass), index = letters[1:4]) %>% 
  mutate(alpha          = rnorm(n = n(), mean = 0, sd = 10),
         beta           = rnorm(n = n(), mean = 0, sd = 1),
         sigma          = runif(n = n(), 0, 10)) %>% 
  mutate(sim_kcal.per.g = rnorm(n = n(), 
                                mean = alpha + beta * log_mass, 
                                sd   = sigma)) %>% 
  
  make_scatter(aes(x = kcal.per.g, y = sim_kcal.per.g))
```

## `b5.7`

The likelihood for `b5.7`, the multivariable `milk` model, is:

$$
\begin{eqnarray}
\text{kcal.per.g}_i & \sim & \text{Normal}(\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta_1 \text{neocortex.perc}_i + \beta_2 \text{log_mass}_i \\
\alpha & \sim & \text{Normal}(0, 100) \\
\beta_1 & \sim & \text{Normal}(0, 1) \\
\beta_2 & \sim & \text{Normal}(0, 1) \\
\sigma & \sim & \text{Uniform}(0, 1)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2, warning = F, message = F}
set.seed(5.7)
d %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, kcal.per.g, neocortex.perc, log_mass), index = letters[1:4]) %>% 
  mutate(alpha          = rnorm(n = n(), mean = 0, sd = 10),
         beta_1         = rnorm(n = n(), mean = 0, sd = 1),
         beta_2         = rnorm(n = n(), mean = 0, sd = 1),
         sigma          = runif(n = n(), 0, 10)) %>% 
  mutate(sim_kcal.per.g = rnorm(n = n(), 
                                mean = alpha + beta_1 * neocortex.perc + beta_2 * log_mass, 
                                sd   = sigma)) %>% 
  
  make_scatter(aes(x = kcal.per.g, y = sim_kcal.per.g))
```

## `b5.8`

For the next few models, let's simulate some leg data.

```{r}
N <- 100
set.seed(531)

d <- 
  tibble(height    = rnorm(N, mean = 10, sd = 2),
         leg_prop  = runif(N, min = 0.4, max = 0.5)) %>% 
  mutate(leg_left  = leg_prop*height + rnorm(N, mean = 0, sd = 0.02),
         leg_right = leg_prop*height + rnorm(N, mean = 0, sd = 0.02))
```

The likelihood for `b5.8` is

$$
\begin{eqnarray}
\text{height}_i & \sim & \text{Normal}(\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta_1 \text{leg_left}_i + \beta_2 \text{leg_right}_i \\
\alpha & \sim & \text{Normal}(10, 100) \\
\beta_1 & \sim & \text{Normal}(2, 10) \\
\beta_2 & \sim & \text{Normal}(2, 10) \\
\sigma & \sim & \text{Uniform}(0, 10)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2}
set.seed(5.8)
d %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, height, leg_left, leg_right), index = letters[1:4]) %>% 
  mutate(alpha      = rnorm(n = n(), mean = 10, sd = 100),
         beta_1     = rnorm(n = n(), mean = 2, sd = 10),
         beta_2     = rnorm(n = n(), mean = 2, sd = 10),
         sigma      = runif(n = n(), 0, 10)) %>% 
  mutate(sim_height = rnorm(n = n(),
                            mean = alpha + beta_1 * leg_left + beta_2 * leg_right,
                            sd   = sigma)) %>% 
  
  make_scatter(aes(x = height, y = sim_height))
```

## `b5.9`

Now we ditch one of the leg lengths, here's the likelihood for the simplified univariable model, `b5.9`

$$
\begin{eqnarray}
\text{height}_i & \sim & \text{Normal}(\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta \text{leg_left}_i \\
\alpha & \sim & \text{Normal}(10, 100) \\
\beta & \sim & \text{Normal}(2, 10) \\
\sigma & \sim & \text{Uniform}(0, 10)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2}
set.seed(5.9)
d %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, height, leg_left), index = letters[1:4]) %>% 
  mutate(alpha      = rnorm(n = n(), mean = 10, sd = 100),
         beta       = rnorm(n = n(), mean = 2, sd = 10),
         sigma      = runif(n = n(), 0, 10)) %>% 
  mutate(sim_height = rnorm(n = n(),
                            mean = alpha + beta * leg_left,
                            sd   = sigma)) %>% 
  
  make_scatter(aes(x = height, y = sim_height))
```

## `b5.10`

Return to the `milk` data.

```{r, message = F}
data(milk)
d <- milk
rm(milk)
```

Here's the likelihood for `b5.10`, the univariable model with `perc.fat`

$$
\begin{eqnarray}
\text{kcal.per.g}_i & \sim & \text{Normal}(\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta \text{perc.fat}_i \\
\alpha & \sim & \text{Normal}(0.6, 10) \\
\beta & \sim & \text{Normal}(0, 1) \\
\sigma & \sim & \text{Uniform}(0, 10)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2}
set.seed(5.10)
d %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, kcal.per.g, perc.fat), index = letters[1:4]) %>% 
  mutate(alpha          = rnorm(n = n(), mean = .6, sd = 10),
         beta           = rnorm(n = n(), mean = 0, sd = 1),
         sigma          = runif(n = n(), 0, 10)) %>% 
  mutate(sim_kcal.per.g = rnorm(n = n(), 
                                mean = alpha + beta * perc.fat, 
                                sd   = sigma)) %>% 
  
  make_scatter(aes(x = kcal.per.g, y = sim_kcal.per.g))
```

## `b5.11`

The likelihood for the alternative univariable model, `b5.11`, is

$$
\begin{eqnarray}
\text{kcal.per.g}_i & \sim & \text{Normal}(\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta \text{perc.lactose}_i \\
\alpha & \sim & \text{Normal}(0.6, 10) \\
\beta & \sim & \text{Normal}(0, 1) \\
\sigma & \sim & \text{Uniform}(0, 10)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2}
set.seed(5.11)
d %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, kcal.per.g, perc.lactose), index = letters[1:4]) %>% 
  mutate(alpha          = rnorm(n = n(), mean = .6, sd = 10),
         beta           = rnorm(n = n(), mean = 0, sd = 1),
         sigma          = runif(n = n(), 0, 10)) %>% 
  mutate(sim_kcal.per.g = rnorm(n = n(), 
                                mean = alpha + beta * perc.lactose, 
                                sd   = sigma)) %>% 
  
  make_scatter(aes(x = kcal.per.g, y = sim_kcal.per.g))
```

## `b5.12`

The likelihood for `b5.12`, the multivariable model, is

$$
\begin{eqnarray}
\text{kcal.per.g}_i & \sim & \text{Normal}(\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta_1 \text{perc.fat}_i + \beta_2 \text{perc.lactose}_i \\
\alpha & \sim & \text{Normal}(0.6, 10) \\
\beta_1 & \sim & \text{Normal}(0, 1) \\
\beta_2 & \sim & \text{Normal}(0, 1) \\
\sigma & \sim & \text{Uniform}(0, 10)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2}
set.seed(5.12)
d %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, kcal.per.g, perc.fat, perc.lactose), index = letters[1:4]) %>% 
  mutate(alpha          = rnorm(n = n(), mean = .6, sd = 10),
         beta_1         = rnorm(n = n(), mean = 0, sd = 1),
         beta_2         = rnorm(n = n(), mean = 0, sd = 1),
         sigma          = runif(n = n(), 0, 10)) %>% 
  mutate(sim_kcal.per.g = rnorm(n = n(), 
                                mean = alpha + beta_1 * perc.fat + beta_2 * perc.lactose, 
                                sd   = sigma)) %>% 
  
  make_scatter(aes(x = kcal.per.g, y = sim_kcal.per.g))
```

## `b5.13`

Let's simulate data, again.

```{r}
N <- 100

set.seed(17)
d <- 
  tibble(h0        = rnorm(N, mean = 10, sd = 2), 
         treatment = rep(0:1, each = N / 2),
         fungus    = rbinom(N, size = 1, prob = .5 - treatment * 0.4),
         h1        = h0 + rnorm(N, mean = 5 - 3 * fungus, sd = 1))
```

The likelihood for the first multivariable `h1` model is

$$
\begin{eqnarray}
\text{h1}_i & \sim & \text{Normal}(\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta_1 \text{h0}_i + \beta_2 \text{treatment}_i + \beta_3 \text{fungus}_i \\
\alpha & \sim & \text{Normal}(0, 100) \\
\beta_1 & \sim & \text{Normal}(0, 10) \\
\beta_2 & \sim & \text{Normal}(0, 10) \\
\beta_3 & \sim & \text{Normal}(0, 10) \\
\sigma & \sim & \text{Uniform}(0, 10)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2}
set.seed(5.13)
d %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, h1, h0, treatment, fungus), index = letters[1:4]) %>% 
  mutate(alpha  = rnorm(n = n(), mean = 0, sd = 100),
         beta_1 = rnorm(n = n(), mean = 0, sd = 10),
         beta_2 = rnorm(n = n(), mean = 0, sd = 10),
         beta_3 = rnorm(n = n(), mean = 0, sd = 10),
         sigma  = runif(n = n(), 0, 10)) %>% 
  mutate(sim_h1 = rnorm(n = n(),
                        mean = alpha + beta_1 * h0 + beta_2 * treatment + beta_3 * fungus,
                        sd   = sigma)) %>% 
  
  make_scatter(aes(x = h1, y = sim_h1))
```

## `b5.14`

The likelihood for the more-reasonable multivariable model is

$$
\begin{eqnarray}
\text{h1}_i & \sim & \text{Normal}(\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta_1 \text{h0}_i + \beta_2 \text{treatment}_i \\
\alpha & \sim & \text{Normal}(0, 100) \\
\beta_1 & \sim & \text{Normal}(0, 10) \\
\beta_2 & \sim & \text{Normal}(0, 10) \\
\sigma & \sim & \text{Uniform}(0, 10)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2}
set.seed(5.14)
d %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, h1, h0, treatment), index = letters[1:4]) %>% 
  mutate(alpha  = rnorm(n = n(), mean = 0, sd = 100),
         beta_1 = rnorm(n = n(), mean = 0, sd = 10),
         beta_2 = rnorm(n = n(), mean = 0, sd = 10),
         sigma  = runif(n = n(), 0, 10)) %>% 
  mutate(sim_h1 = rnorm(n = n(),
                        mean = alpha + beta_1 * h0 + beta_2 * treatment,
                        sd   = sigma)) %>% 
  
  make_scatter(aes(x = h1, y = sim_h1))
```

## `b5.15`

Now load the `Howell1` data.

```{r, message = F}
data(Howell1)
d <- Howell1
rm(Howell1)
```

The likelihood for `5.15`, the univariable `height` model, is

$$
\begin{eqnarray}
\text{height}_i & \sim & \text{Normal}(\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta \text{male}_i \\
\alpha & \sim & \text{Normal}(178, 100) \\
\beta & \sim & \text{Normal}(0, 10) \\
\sigma & \sim & \text{HalfCauchy}(0, 2)
\end{eqnarray}
$$


```{r, fig.width = 8, fig.height = 2}
set.seed(5.15)
d %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, height, male), index = letters[1:4]) %>% 
  mutate(alpha      = rnorm(n = n(), mean = 178, sd = 100),
         beta       = rnorm(n = n(), mean = 0, sd = 10),
         sigma      = rcauchy(n = n(), location = 0, scale = 2) %>% abs()) %>% 
  mutate(sim_height = rnorm(n = n(),
                            mean = alpha + beta * male,
                            sd   = sigma)) %>% 
  
  make_scatter(aes(x = height, y = sim_height))
```

## `b5.15b`

Here's the likelihood for the reparameterized `b5.15b` model

$$
\begin{eqnarray}
\text{height}_i & \sim & \text{Normal}(\mu_i, \sigma) \\
\mu_i & = & \beta_1 \text{male}_i + \beta_2 \text{female}_i \\
\beta_1 & \sim & \text{Normal}(178, 100) \\
\beta_2 & \sim & \text{Normal}(178, 100) \\
\sigma & \sim & \text{HalfCauchy}(0, 2)
\end{eqnarray}
$$

The model requires us to make a `female` dummy.

```{r}
d <-
  d %>%
  mutate(female = 1 - male)
```


```{r, fig.width = 8, fig.height = 2}
set.seed(5.15)
d %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, height, male, female), index = letters[1:4]) %>% 
  mutate(beta_1     = rnorm(n = n(), mean = 178, sd = 100),
         beta_2     = rnorm(n = n(), mean = 178, sd = 100),
         sigma      = rcauchy(n = n(), location = 0, scale = 2) %>% abs()) %>% 
  mutate(sim_height = rnorm(n = n(),
                            mean =  beta_1 * male + beta_2 * female,
                            sd   = sigma)) %>% 
  
  make_scatter(aes(x = height, y = sim_height))
```

## `b5.16`

Reload the `milk` data.

```{r, message = F}
data(milk)
d <- milk
rm(milk)
```

As `clade` has 4 categories, let's convert these to 4 dummy variables.

```{r}
d <- 
  d %>%
  mutate(clade_nwm = ifelse(clade == "New World Monkey", 1, 0),
         clade_owm = ifelse(clade == "Old World Monkey", 1, 0),
         clade_s   = ifelse(clade == "Strepsirrhine", 1, 0),
         clade_ape = ifelse(clade == "Ape", 1, 0))
```

Here's the likelihood using three of those dummies

$$
\begin{eqnarray}
\text{kcal.per.g}_i & \sim & \text{Normal}(\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta_1 \text{clade_nwm}_i + \beta_2 \text{clade_owm}_i + \beta_3 \text{clade_s}_i  \\
\alpha & \sim & \text{Normal}(0.6, 10) \\
\beta_1 & \sim & \text{Normal}(0, 1) \\
\beta_2 & \sim & \text{Normal}(0, 1) \\
\beta_3 & \sim & \text{Normal}(0, 1) \\
\sigma & \sim & \text{Uniform}(0, 10)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2}
set.seed(5.16)
d %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, kcal.per.g, clade_nwm, clade_owm, clade_s), index = letters[1:4]) %>% 
  mutate(alpha          = rnorm(n = n(), mean = 0.6, sd = 10),
         beta_1         = rnorm(n = n(), mean = 0, sd = 1),
         beta_2         = rnorm(n = n(), mean = 0, sd = 1),
         beta_3         = rnorm(n = n(), mean = 0, sd = 1),
         sigma          = runif(n = n(), 0, 10)) %>% 
  mutate(sim_kcal.per.g = rnorm(n = n(), 
                                mean = alpha + beta_1 * clade_nwm + beta_2 * clade_owm + beta_3 * clade_s, 
                                sd   = sigma)) %>% 
  
  make_scatter(aes(x = kcal.per.g, y = sim_kcal.per.g))
```

## `b5.17`

The likelihood for the alternative parameterization is

$$
\begin{eqnarray}
\text{kcal.per.g}_i & \sim & \text{Normal}(\mu_i, \sigma) \\
\mu_i & = & \beta_1 \text{clade_nwm}_i + \beta_2 \text{clade_owm}_i + \beta_3 \text{clade_s}_i + \beta_4 \text{clade_ape}_i  \\
\beta_1 & \sim & \text{Normal}(0.6, 10) \\
\beta_2 & \sim & \text{Normal}(0.6, 10) \\
\beta_3 & \sim & \text{Normal}(0.6, 10) \\
\beta_4 & \sim & \text{Normal}(0.6, 10) \\
\sigma & \sim & \text{Uniform}(0, 10)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2}
set.seed(5.17)
d %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, kcal.per.g, clade_nwm, clade_owm, clade_s, clade_ape), index = letters[1:4]) %>% 
  mutate(beta_1         = rnorm(n = n(), mean = 0.6, sd = 10),
         beta_2         = rnorm(n = n(), mean = 0.6, sd = 10),
         beta_3         = rnorm(n = n(), mean = 0.6, sd = 10),
         beta_4         = rnorm(n = n(), mean = 0.6, sd = 10),
         sigma          = runif(n = n(), 0, 10)) %>% 
  mutate(sim_kcal.per.g = rnorm(n = n(), 
                                mean = beta_1 * clade_nwm + beta_2 * clade_owm + beta_3 * clade_s + beta_4 * clade_ape, 
                                sd   = sigma)) %>% 
  
  make_scatter(aes(x = kcal.per.g, y = sim_kcal.per.g))
```

## Reference {-}

[McElreath, R. (2016). *Statistical rethinking: A Bayesian course with examples in R and Stan.* Chapman & Hall/CRC Press.](https://xcelab.net/rm/statistical-rethinking/)

## Session info {-}

```{r}
sessionInfo()
```

```{r, warning = F, echo = F}
rm(d, N, make_scatter)
```

