---
title: "Ch. 6 Overfitting, Regularization, and Information Criteria"
author: "A Solomon Kurz"
date: "`r format(Sys.Date())`"
output:
  html_document
---

```{r set-options, echo = F, cache = F}
options(width = 100)
```

Let's start of by making the data with brain size and body size for seven `species`.

```{r, warning = F, message = F}
library(tidyverse)

(
  d <- 
  tibble(species = c("afarensis", "africanus", "habilis", "boisei", "rudolfensis", "ergaster", "sapiens"), 
         brain   = c(438, 452, 612, 521, 752, 871, 1350), 
         mass    = c(37.0, 35.5, 34.5, 41.5, 55.5, 61.0, 53.5)) %>%
  mutate(mass_s  = (mass - mean(mass)) / sd(mass))
  )
```

Models `b6.1` through `b6.7` were fit with the frequentist `lm()`. Since there are no prior distributions for those models, they leave us on prior predictive distributions to plot. 

## `b6.8`

Here's the likelihood for `b6.8`:

$$
\begin{eqnarray}
\text{brain}_i & \sim & \text{Normal}(\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta \text{mass_s}_i \\
\alpha & \sim & \text{Normal}(0, 1000) \\
\beta & \sim & \text{Normal}(0, 1000) \\
\sigma & \sim & \text{HalfCauchy}(0, 10)
\end{eqnarray}
$$

Before we plot, let's make our custom plotting function, `make_scatter()`.

```{r}
make_scatter <- function(data, mapping, ...){
  ggplot(data, mapping) +
  geom_abline(color = "white") +
  geom_point(alpha = 1/2, size = 1/2) +
  theme(panel.grid = element_blank()) +
  facet_wrap(~index, scales = "free_y", ncol = 4)
}
```

```{r, fig.width = 8, fig.height = 2}
set.seed(6.8)
d %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, brain, mass_s), index = letters[1:4]) %>% 
  mutate(alpha     = rnorm(n = n(), mean = 0, sd = 1000),
         beta      = rnorm(n = n(), mean = 0, sd = 1000),
         sigma     = rcauchy(n = n(), location = 0, scale = 10) %>% abs()) %>% 
  mutate(sim_brain = rnorm(n = n(), 
                           mean = alpha + beta * mass_s, 
                           sd   = sigma)) %>% 
  
  make_scatter(aes(x = brain, y = sim_brain))
```




Load the `milk` data from earlier in the text.

```{r, message = F, warning = F}
library(rethinking)

data(milk)
d <- 
  milk %>%
  filter(complete.cases(.)) %>%
  mutate(neocortex = neocortex.perc / 100)

rm(milk)
```

The likelihood for model `b6.11` is 

$$
\begin{eqnarray}
\text{kcal.per.g}_i & \sim & \text{Normal}(\mu, \sigma) \\
\mu & \sim & \text{Uniform}(-1000, 1000) \\
\sigma & \sim & \text{Uniform}(0, 100)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2}
set.seed(6.11)
d %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, kcal.per.g), index = letters[1:4]) %>% 
  mutate(alpha          = runif(n = n(), -1000, 1000),
         sigma          = runif(n = n(), 0, 100)) %>% 
  mutate(sim_kcal.per.g = rnorm(n = n(), mean = alpha, sd = sigma)) %>% 
  
  make_scatter(aes(x = kcal.per.g, y = sim_kcal.per.g))
```

Here's the likelihood for `b6.12`, the univariable model with `neocortex` as the predictor:

$$
\begin{eqnarray}
\text{kcal.per.g}_i & \sim & \text{Normal}(\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta \text{neocortex}_i \\
\alpha & \sim & \text{Uniform}(-1000, 1000) \\
\beta & \sim & \text{Uniform}(-1000, 1000) \\
\sigma & \sim & \text{Uniform}(0, 100)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2}
set.seed(6.12)
d %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, kcal.per.g, neocortex), index = letters[1:4]) %>% 
  mutate(alpha          = runif(n = n(), -1000, 1000),
         beta           = runif(n = n(), -1000, 1000),
         sigma          = runif(n = n(), 0, 100)) %>% 
  mutate(sim_kcal.per.g = rnorm(n = n(), 
                                mean = alpha + beta * neocortex, 
                                sd   = sigma)) %>% 
  
  make_scatter(aes(x = kcal.per.g, y = sim_kcal.per.g))
```

Now here's the likelihood for `b6.13`, the univariable model with `log(mass)` as the predictor:

$$
\begin{eqnarray}
\text{kcal.per.g}_i & \sim & \text{Normal}(\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta \text{log(mass)}_i \\
\alpha & \sim & \text{Uniform}(-1000, 1000) \\
\beta & \sim & \text{Uniform}(-1000, 1000) \\
\sigma & \sim & \text{Uniform}(0, 100)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2}
set.seed(6.13)
d %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, kcal.per.g, mass), index = letters[1:4]) %>% 
  mutate(alpha          = runif(n = n(), -1000, 1000),
         beta           = runif(n = n(), -1000, 1000),
         sigma          = runif(n = n(), 0, 100)) %>% 
  mutate(sim_kcal.per.g = rnorm(n = n(), 
                                mean = alpha + beta * log(mass), 
                                sd   = sigma)) %>% 
  
  make_scatter(aes(x = kcal.per.g, y = sim_kcal.per.g))
```

Finally, here's the multivariable likelihood, model `b6.14`:

$$
\begin{eqnarray}
\text{kcal.per.g}_i & \sim & \text{Normal}(\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta_1 \text{neocortex}_i + \beta_2 \text{log(mass)}_i \\
\alpha & \sim & \text{Uniform}(-1000, 1000) \\
\beta_1 & \sim & \text{Uniform}(-1000, 1000) \\
\beta_2 & \sim & \text{Uniform}(-1000, 1000) \\
\sigma & \sim & \text{Uniform}(0, 100)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2}
set.seed(6.14)
d %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, kcal.per.g, neocortex, mass), index = letters[1:4]) %>% 
  mutate(alpha          = runif(n = n(), -1000, 1000),
         beta_1         = runif(n = n(), -1000, 1000),
         beta_2         = runif(n = n(), -1000, 1000),
         sigma          = runif(n = n(), 0, 100)) %>% 
  mutate(sim_kcal.per.g = rnorm(n = n(), 
                                mean = alpha + beta_1 * neocortex + beta_2 * log(mass), 
                                sd   = sigma)) %>% 
  
  make_scatter(aes(x = kcal.per.g, y = sim_kcal.per.g))
```

## Reference {-}

[McElreath, R. (2016). *Statistical rethinking: A Bayesian course with examples in R and Stan.* Chapman & Hall/CRC Press.](https://xcelab.net/rm/statistical-rethinking/)

## Session info {-}

```{r}
sessionInfo()
```

```{r, warning = F, echo = F}
rm(d, make_scatter)
```

