---
title: "Ch. 7 Interactions"
author: "A Solomon Kurz"
date: "`r format(Sys.Date())`"
output:
  html_document
---

```{r set-options, echo = F, cache = F}
options(width = 100)
```

Load the rethinking package and the `rugged` data.

```{r, message = F, warning = F}
library(rethinking)

data(rugged)
d <- rugged
rm(rugged)
```

Now load the tidyverse and wrangle the data a bit.

```{r, message = F, warning = F}
library(tidyverse)

# make the log version of criterion
d <- 
  d %>%
  mutate(log_gdp = log(rgdppc_2000))

# extract countries with GDP data
dd <-
  d %>%
  filter(complete.cases(rgdppc_2000))

# split countries into Africa and not-Africa
d.A1 <-
  dd %>%
  filter(cont_africa == 1)

d.A0 <-
  dd %>%
  filter(cont_africa == 0)
```

## `b7.1`

The likelihood for the univariable model with the `d.A1` data is

$$
\begin{eqnarray}
\text{log_gdp}_{\text{data = d.A1}_i} & \sim & \text{Normal} (\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta \text{rugged}_i \\
\alpha & \sim & \text{Normal}(8, 100) \\
\beta & \sim & \text{Normal}(0, 1) \\
\sigma & \sim & \text{Uniform}(0, 10)
\end{eqnarray}
$$

Before we plot, let's make our custom plotting function, `make_scatter()`.

```{r}
make_scatter <- function(data, mapping, ...){
  ggplot(data, mapping) +
  geom_abline(color = "white") +
  geom_point(alpha = 1/2, size = 1/2) +
  theme(panel.grid = element_blank()) +
  facet_wrap(~index, scales = "free_y", ncol = 4)
}
```

```{r, fig.width = 8, fig.height = 2}
set.seed(7.1)
d.A1 %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, log_gdp, rugged), index = letters[1:4]) %>% 
  mutate(alpha       = rnorm(n = n(), mean = 8, sd = 100),
         beta        = rnorm(n = n(), mean = 0, sd = 1),
         sigma       = runif(n = n(), 0, 10)) %>% 
  mutate(sim_log_gdp = rnorm(n = n(),
                             mean = alpha + beta * rugged, 
                             sd   = sigma)) %>% 
  
  make_scatter(aes(x = log_gdp, y = sim_log_gdp))
```

## `b7.2`

The likelihood for the next univariable model, `b7.2`, is very much the same as the last. The only difference is this time we use the `d.A0` data.

$$
\begin{eqnarray}
\text{log_gdp}_{\text{data = d.A0}_i} & \sim & \text{Normal} (\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta \text{rugged}_i \\
\alpha & \sim & \text{Normal}(8, 100) \\
\beta & \sim & \text{Normal}(0, 1) \\
\sigma & \sim & \text{Uniform}(0, 10)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2}
set.seed(7.2)
d.A0 %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, log_gdp, rugged), index = letters[1:4]) %>% 
  mutate(alpha       = rnorm(n = n(), mean = 8, sd = 100),
         beta        = rnorm(n = n(), mean = 0, sd = 1),
         sigma       = runif(n = n(), 0, 10)) %>% 
  mutate(sim_log_gdp = rnorm(n = n(),
                             mean = alpha + beta * rugged, 
                             sd   = sigma)) %>% 
  
  make_scatter(aes(x = log_gdp, y = sim_log_gdp))
```

## `b7.3`

The likelihood for `b7.3` followed the same trend, this time with the `dd` data.

$$
\begin{eqnarray}
\text{log_gdp}_{\text{data = dd}_i} & \sim & \text{Normal} (\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta \text{rugged}_i \\
\alpha & \sim & \text{Normal}(8, 100) \\
\beta & \sim & \text{Normal}(0, 1) \\
\sigma & \sim & \text{Uniform}(0, 10)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2}
set.seed(7.3)
dd %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, log_gdp, rugged), index = letters[1:4]) %>% 
  mutate(alpha       = rnorm(n = n(), mean = 8, sd = 100),
         beta        = rnorm(n = n(), mean = 0, sd = 1),
         sigma       = runif(n = n(), 0, 10)) %>% 
  mutate(sim_log_gdp = rnorm(n = n(),
                             mean = alpha + beta * rugged, 
                             sd   = sigma)) %>% 
  
  make_scatter(aes(x = log_gdp, y = sim_log_gdp))
```

## `b7.4`

With model `b7.4` we finally switch the likelihood up a bit.

$$
\begin{eqnarray}
\text{log_gdp}_{\text{data = dd}_i} & \sim & \text{Normal} (\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta_1 \text{rugged}_i + \beta_2 \text{cont_africa}_i \\
\alpha & \sim & \text{Normal}(8, 100) \\
\beta_1 & \sim & \text{Normal}(0, 1) \\
\beta_2 & \sim & \text{Normal}(0, 1) \\
\sigma & \sim & \text{Uniform}(0, 10)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2}
set.seed(7.4)
dd %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, log_gdp, rugged, cont_africa), index = letters[1:4]) %>% 
  mutate(alpha       = rnorm(n = n(), mean = 8, sd = 100),
         beta_1      = rnorm(n = n(), mean = 0, sd = 1),
         beta_2      = rnorm(n = n(), mean = 0, sd = 1),
         sigma       = runif(n = n(), 0, 10)) %>% 
  mutate(sim_log_gdp = rnorm(n = n(),
                             mean = alpha + beta_1 * rugged + beta_2 * cont_africa, 
                             sd   = sigma)) %>% 
  
  make_scatter(aes(x = log_gdp, y = sim_log_gdp))
```

## `b7.5`

The likelihood for `b7.5` follows the form

$$
\begin{eqnarray}
\text{log_gdp}_{\text{data = dd}_i} & \sim & \text{Normal} (\mu_i, \sigma) \\
\mu_i & = & \alpha + \gamma_i \text{rugged}_i + \beta_2 \text{cont_africa}_i \\
\gamma_i & = & \beta_1 + \beta_3 \text{cont_africa}_i \\
\alpha & \sim & \text{Normal}(8, 100) \\
\beta_1 & \sim & \text{Normal}(0, 1) \\
\beta_2 & \sim & \text{Normal}(0, 1) \\
\beta_3 & \sim & \text{Normal}(0, 1) \\
\sigma & \sim & \text{Uniform}(0, 10)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2}
set.seed(7.5)
dd %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, log_gdp, rugged, cont_africa), index = letters[1:4]) %>% 
  mutate(alpha       = rnorm(n = n(), mean = 8, sd = 100),
         beta_1      = rnorm(n = n(), mean = 0, sd = 1),
         beta_2      = rnorm(n = n(), mean = 0, sd = 1),
         beta_3      = rnorm(n = n(), mean = 0, sd = 1),
         sigma       = runif(n = n(), 0, 10)) %>% 
  mutate(gamma       = beta_1 + beta_3 * cont_africa) %>% 
  mutate(sim_log_gdp = rnorm(n = n(),
                             mean = alpha + gamma * rugged + beta_2 * cont_africa, 
                             sd   = sigma)) %>% 
  
  make_scatter(aes(x = log_gdp, y = sim_log_gdp))
```

## `b7.5b`

The likelihood for `b7.5b` follows the more-conventional form

$$
\begin{eqnarray}
\text{log_gdp}_i & \sim & \text{Normal} (\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta_1 \text{rugged}_i + \beta_2 \text{cont_africa}_i + \beta_3 \text{rugged}_i \times \text{cont_africa}_i \\
\alpha & \sim & \text{Normal}(8, 100) \\
\beta_1 & \sim & \text{Normal}(0, 1) \\
\beta_2 & \sim & \text{Normal}(0, 1) \\
\beta_3 & \sim & \text{Normal}(0, 1) \\
\sigma & \sim & \text{Uniform}(0, 10)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2}
set.seed(7.5)
dd %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, log_gdp, rugged, cont_africa), index = letters[1:4]) %>% 
  mutate(alpha       = rnorm(n = n(), mean = 8, sd = 100),
         beta_1      = rnorm(n = n(), mean = 0, sd = 1),
         beta_2      = rnorm(n = n(), mean = 0, sd = 1),
         beta_3      = rnorm(n = n(), mean = 0, sd = 1),
         sigma       = runif(n = n(), 0, 10)) %>% 
  mutate(sim_log_gdp = rnorm(n = n(),
                             mean = alpha + beta_1 * rugged + beta_2 * cont_africa + beta_3 * rugged * cont_africa, 
                             sd   = sigma)) %>% 
  
  make_scatter(aes(x = log_gdp, y = sim_log_gdp))
```

## `b7.6`

The next model requires the `tulips` data.

```{r}
data(tulips)
d <- tulips
rm(tulips)
```

The likelihood for `b7.6` follows the form

$$
\begin{eqnarray}
\text{blooms}_i & \sim & \text{Normal} (\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta_1 \text{water}_i + \beta_2 \text{shade}_i \\
\alpha & \sim & \text{Normal}(0, 100) \\
\beta_1 & \sim & \text{Normal}(0, 100) \\
\beta_2 & \sim & \text{Normal}(0, 100) \\
\sigma & \sim & \text{Uniform}(0, 100)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2}
set.seed(7.6)
d %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, blooms, water, shade), index = letters[1:4]) %>% 
  mutate(alpha      = rnorm(n = n(), mean = 0, sd = 100),
         beta_1     = rnorm(n = n(), mean = 0, sd = 100),
         beta_2     = rnorm(n = n(), mean = 0, sd = 100),
         sigma      = runif(n = n(), 0, 100)) %>% 
  mutate(sim_blooms = rnorm(n = n(),
                             mean = alpha + beta_1 * water + beta_2 * shade, 
                             sd   = sigma)) %>% 
  
  make_scatter(aes(x = blooms, y = sim_blooms))
```

## `b7.7`

The likelihood for model `b7.7` added an interaction term.

$$
\begin{eqnarray}
\text{blooms}_i & \sim & \text{Normal} (\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta_1 \text{water}_i + \beta_2 \text{shade}_i + \beta_3 \text{water}_i \times \text{shade}_i \\
\alpha & \sim & \text{Normal}(0, 100) \\
\beta_1 & \sim & \text{Normal}(0, 100) \\
\beta_2 & \sim & \text{Normal}(0, 100) \\
\beta_3 & \sim & \text{Normal}(0, 100) \\
\sigma & \sim & \text{Uniform}(0, 100)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2}
set.seed(7.7)
d %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, blooms, water, shade), index = letters[1:4]) %>% 
  mutate(alpha      = rnorm(n = n(), mean = 0, sd = 100),
         beta_1     = rnorm(n = n(), mean = 0, sd = 100),
         beta_2     = rnorm(n = n(), mean = 0, sd = 100),
         beta_3     = rnorm(n = n(), mean = 0, sd = 100),
         sigma      = runif(n = n(), 0, 100)) %>% 
  mutate(sim_blooms = rnorm(n = n(),
                             mean = alpha + beta_1 * water + beta_2 * shade + beta_3 * water * shade, 
                             sd   = sigma)) %>% 
  
  make_scatter(aes(x = blooms, y = sim_blooms))
```

## `b7.8`

The next two likelihoods require us to center the predictors.

```{r}
d <-
  d %>%
  mutate(shade_c = shade - mean(shade),
         water_c = water - mean(water))
```

Now the likelihood for `b7.8` follows the form

$$
\begin{eqnarray}
\text{blooms}_i & \sim & \text{Normal} (\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta_1 \text{water_c}_i + \beta_2 \text{shade_c}_i \\
\alpha & \sim & \text{Normal}(130, 100) \\
\beta_1 & \sim & \text{Normal}(0, 100) \\
\beta_2 & \sim & \text{Normal}(0, 100) \\
\sigma & \sim & \text{HalfCauchy}(0, 10)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2}
set.seed(7.8)
d %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, blooms, water_c, shade_c), index = letters[1:4]) %>% 
  mutate(alpha      = rnorm(n = n(), mean = 130, sd = 100),
         beta_1     = rnorm(n = n(), mean = 0, sd = 100),
         beta_2     = rnorm(n = n(), mean = 0, sd = 100),
         sigma      = rcauchy(n = n(), location = 0, scale = 10) %>% abs()) %>% 
  mutate(sim_blooms = rnorm(n = n(),
                             mean = alpha + beta_1 * water_c + beta_2 * shade_c, 
                             sd   = sigma)) %>% 
  
  make_scatter(aes(x = blooms, y = sim_blooms))
```

## `b7.9`

The likelihood for `b7.9` follows the form

$$
\begin{eqnarray}
\text{blooms}_i & \sim & \text{Normal} (\mu_i, \sigma) \\
\mu_i & = & \alpha + \beta_1 \text{water_c}_i + \beta_2 \text{shade_c}_i  + \beta_3 \text{water_c}_i \times \text{shade_c}_i \\
\alpha & \sim & \text{Normal}(130, 100) \\
\beta_1 & \sim & \text{Normal}(0, 100) \\
\beta_2 & \sim & \text{Normal}(0, 100) \\
\beta_3 & \sim & \text{Normal}(0, 100) \\
\sigma & \sim & \text{HalfCauchy}(0, 10)
\end{eqnarray}
$$

```{r, fig.width = 8, fig.height = 2}
set.seed(7.9)
d %>% 
  mutate(row = 1:n()) %>% 
  expand(nesting(row, blooms, water_c, shade_c), index = letters[1:4]) %>% 
  mutate(alpha      = rnorm(n = n(), mean = 130, sd = 100),
         beta_1     = rnorm(n = n(), mean = 0, sd = 100),
         beta_2     = rnorm(n = n(), mean = 0, sd = 100),
         beta_3     = rnorm(n = n(), mean = 0, sd = 100),
         sigma      = rcauchy(n = n(), location = 0, scale = 10) %>% abs()) %>% 
  mutate(sim_blooms = rnorm(n = n(),
                             mean = alpha + beta_1 * water_c + beta_2 * shade_c + beta_3 * water_c * shade_c, 
                             sd   = sigma)) %>% 
  
  make_scatter(aes(x = blooms, y = sim_blooms))
```

## Reference {-}

[McElreath, R. (2016). *Statistical rethinking: A Bayesian course with examples in R and Stan.* Chapman & Hall/CRC Press.](https://xcelab.net/rm/statistical-rethinking/)

## Session info {-}

```{r}
sessionInfo()
```

```{r, warning = F, echo = F}
rm(d, d.A0, d.A1, dd, make_scatter)
```

