---
title: "Section 12.4.1. Posterior prediction for same clusters"
author: "A Solomon Kurz"
date: "`r format(Sys.Date())`"
output:
  github_document
---

```{r, echo = FALSE, cache = FALSE}
knitr::opts_chunk$set(fig.retina = 2.5)
options(width = 100)
```

These packages will already have been loaded.

```{r, warning = F, message = F}
library(tidyverse)
library(brms)
library(ggthemes) 
library(patchwork)
library(tidybayes)
library(bayesplot)
library(posterior)
```

Now load the focal model, `b12.4`.

```{r}
b12.4 <- readRDS("/Users/solomonkurz/Dropbox/Recoding McElreath/fits/b12.04.rds")
```

Review the model.

```{r}
print(b12.4)
```


## 12.4 Multilevel posterior predictions

### 12.4.1 Posterior prediction for same clusters.

```{r}
post <- posterior_samples(b12.4)

glimpse(post)
```

### 12.4.2 Posterior prediction for new clusters.

By average actor, McElreath referred to a chimp with an intercept exactly at the population mean $\alpha$. So this time we'll only be working with the population parameters, or what are also sometimes called the fixed effects. When using `brms::posterior_samples()` output, this would mean working with columns beginning with the `b_` prefix (i.e., `b_Intercept`, `b_prosoc_left`, and `b_prosoc_left:condition`).

```{r, message = F}
post_average_actor <-
  post %>% 
  # here we use the linear regression formula to get the log_odds for the 4 conditions
  transmute(`0/0` = b_Intercept,
            `1/0` = b_Intercept + b_prosoc_left,
            `0/1` = b_Intercept,
            `1/1` = b_Intercept + b_prosoc_left + `b_prosoc_left:condition`) %>%
  # with `mutate_all()` we can convert the estimates to probabilities in one fell swoop
  mutate_all(inv_logit_scaled) %>% 
  # putting the data in the long format and grouping by condition (i.e., `key`)
  gather() %>%
  mutate(key = factor(key, level = c("0/0", "1/0", "0/1", "1/1"))) %>% 
  group_by(key) %>%
  # here we get the summary values for the plot
  summarise(m  = mean(value),
            # note we're using 80% intervals
            ll = quantile(value, probs = .1),
            ul = quantile(value, probs = .9))

post_average_actor
```

Figure 12.5.a.

```{r, fig.width = 2.5, fig.height = 3.25}
p1 <-
  post_average_actor %>%
  ggplot(aes(x = key, y = m, group = 1)) +
  geom_ribbon(aes(ymin = ll, ymax = ul), fill = "orange1") +
  geom_line(color = "blue") +
  ggtitle("Average actor") +
  coord_cartesian(ylim = c(0, 1)) +
  theme_fivethirtyeight() +
  theme(plot.title = element_text(size = 14, hjust = .5))

p1
```

If we want to depict the variability across the chimps, we need to include `sd_actor__Intercept` into the calculations. In the first block of code, below, we simulate a bundle of new intercepts defined by

$$\alpha_\text{actor} \sim \operatorname{Normal}(0, \sigma_\text{actor}).$$

```{r}
# the random effects
set.seed(12.42)
ran_ef <-
  tibble(random_effect = rnorm(n = 1000, mean = 0, sd = post$sd_actor__Intercept))

glimpse(ran_ef)
```

To get a sense of what we've done, we'll plot these with a histogram.

```{r, fig.width = 4, fig.height = 2.75}
ran_ef %>% 
  
  ggplot(aes(x = random_effect)) +
  geom_histogram(fill = "blue", binwidth = 1) +
  scale_y_continuous(breaks = NULL) +
  ggtitle("1,000 devations, representing\nnew actor effects") +
  theme_fivethirtyeight()
```

Since they are deviations, they center around zero.


```{r}
fix_ef <-
  post %>% 
  slice(1:1000) %>%
  transmute(`0/0` = b_Intercept,
            `1/0` = b_Intercept + b_prosoc_left,
            `0/1` = b_Intercept,
            `1/1` = b_Intercept + b_prosoc_left + `b_prosoc_left:condition`)

fix_ef
```


```{r}
bind_cols(ran_ef, fix_ef) %>% 
  gather(-random_effect, key = "condition", value = "fixed_effect") %>% 
  mutate(condition = factor(condition, level = c("0/0", "1/0", "0/1", "1/1"))) %>%
  mutate(intercept = fixed_effect + random_effect) %>%
  mutate(prob = inv_logit_scaled(intercept)) %>%
  group_by(condition) %>%
  summarise(m  = mean(prob),
            ll = quantile(prob, probs = .1),
            ul = quantile(prob, probs = .9))
```


new version

```{r, message = F}
# arrange the fixed effects (i.e., the population parameters)
fix_ef <-
  post %>% 
  slice(1:1000) %>%
  transmute(`0/0` = b_Intercept,
            `1/0` = b_Intercept + b_prosoc_left,
            `0/1` = b_Intercept,
            `1/1` = b_Intercept + b_prosoc_left + `b_prosoc_left:condition`)

# combine `ran_ef` and `fix_ef`
ran_and_fix_ef <-
  bind_cols(ran_ef, fix_ef) %>% 
  gather(-random_effect, key = "condition", value = "fixed_effect") %>% 
  mutate(condition = factor(condition, level = c("0/0", "1/0", "0/1", "1/1"))) %>%
  mutate(intercept = fixed_effect + random_effect) %>%
  mutate(prob = inv_logit_scaled(intercept))

# to simplify things, we'll reduce them to summaries
(
  marginal_effects <-
  ran_and_fix_ef %>%
  group_by(condition) %>%
  summarise(m  = mean(prob),
            ll = quantile(prob, probs = .1),
            ul = quantile(prob, probs = .9))
)
```

Behold Figure 12.5.b.

```{r, fig.width = 2.5, fig.height = 3.25}
p2 <-
  marginal_effects %>%
  ggplot(aes(x = condition, y = m, group = 1)) +
  geom_ribbon(aes(ymin = ll, ymax = ul), fill = "orange1") +
  geom_line(color = "blue") +
  ggtitle("Marginal of actor") +
  coord_cartesian(ylim = c(0, 1)) +
  theme_fivethirtyeight() +
  theme(plot.title = element_text(size = 14, hjust = .5))

p2
```

Figure 12.5.c just takes a tiny bit more wrangling. 

```{r, fig.width = 2.5, fig.height = 3.25}
p3 <-
  ran_and_fix_ef %>%
  mutate(iter = rep(1:1000, times = 4)) %>%
  filter(iter %in% c(1:50)) %>%
  
  ggplot(aes(x = condition, y = prob, group = iter)) +
  geom_line(alpha = 1/2, color = "orange3") +
  ggtitle("50 simulated actors") +
  coord_cartesian(ylim = c(0, 1)) +
  theme_fivethirtyeight() +
  theme(plot.title = element_text(size = 14, hjust = .5))

p3
```

For the finale, we'll stitch the three plots together.

```{r, fig.width = 7.5, fig.height = 3.25, message = F, warning = F}
p1 | p2 | p3
```


#### Bonus: Let's use `fitted()` this time. 

We just made those plots using various wrangled versions of `post`, the data frame returned by `posterior_samples(b.12.4)`. If you followed along closely, part of what made that a great exercise is that it forced you to consider what the various vectors in `post` meant with respect to the model formula. But it's also handy to see how to do that from a different perspective. So in this section, we'll repeat that process by relying on the `fitted()` function, instead. We'll go in the same order, starting with the average actor.

```{r, warning = F, message = F}
nd <-
  tibble(prosoc_left = c(0, 1, 0, 1),
         condition   = c(0, 0, 1, 1))

(
  f <-
  fitted(b12.4,
         newdata = nd,
         re_formula = NA,
         probs = c(.1, .9)) %>% 
  as_tibble() %>% 
  bind_cols(nd) %>% 
  mutate(condition = factor(str_c(prosoc_left, "/", condition), 
                            levels = c("0/0", "1/0", "0/1", "1/1")))
)
```

You should notice a few things. Since `b12.4` is a multilevel model, it had three predictors: `prosoc_left`, `condition`, and `actor`. However, our `nd` data only included the first two of those predictors. The reason `fitted()` permitted that was because we set `re_formula = NA`. When you do that, you tell `fitted()` to ignore group-level effects (i.e., focus only on the fixed effects). This was our `fitted()` version of ignoring the `r_` vectors returned by `posterior_samples()`. Here's the plot.

```{r, fig.width = 2.5, fig.height = 3.25}
p4 <-
  f %>%
  ggplot(aes(x = condition, y = Estimate, group = 1)) +
  geom_ribbon(aes(ymin = Q10, ymax = Q90), fill = "blue") +
  geom_line(color = "orange1") +
  ggtitle("Average actor") +
  coord_cartesian(ylim = c(0, 1)) +
  theme_fivethirtyeight() +
  theme(plot.title = element_text(size = 14, hjust = .5))

p4
```

For marginal of actor, we can continue using the same `nd` data. This time we'll be sticking with the default `re_formula` setting, which will accommodate the multilevel nature of the model. However, we'll also be adding `allow_new_levels = T` and `sample_new_levels = "gaussian"`. The former will allow us to marginalize across the specific actors in our data and the latter will instruct `fitted()` to use the multivariate normal distribution implied by the random effects. It'll make more sense why I say *multivariate* normal by the end of the [next chapter][Adventures in Covariance]. For now, just go with it.

```{r, warning = F, message = F}
(
  f <-
  fitted(b12.4,
         newdata = nd,
         probs = c(.1, .9),
         allow_new_levels = T,
         sample_new_levels = "gaussian") %>% 
  as_tibble() %>% 
  bind_cols(nd) %>% 
  mutate(condition = factor(str_c(prosoc_left, "/", condition), 
                            levels = c("0/0", "1/0", "0/1", "1/1")))
  )
```

Here's our `fitted()`-based marginal of `actor` plot.

```{r, fig.width = 2.5, fig.height = 3.25}
p5 <-
  f %>%
  ggplot(aes(x = condition, y = Estimate, group = 1)) +
  geom_ribbon(aes(ymin = Q10, ymax = Q90), fill = "blue") +
  geom_line(color = "orange1") +
  ggtitle("Marginal of actor") +
  coord_cartesian(ylim = c(0, 1)) +
  theme_fivethirtyeight() +
  theme(plot.title = element_text(size = 14, hjust = .5))

p5
```

```{r}
nd <-
  tibble(prosoc_left = c(0, 1, 0, 1),
         condition   = c(0, 0, 1, 1)) %>% 
  expand(actor = c(str_c("new0", 1:9), str_c("new", 10:50)),
         nesting(prosoc_left, condition)) %>% 
  # this adds a row number, which will come in handy, later
  mutate(row = 1:n())

glimpse(nd)
```


```{r}
nd <-
  tibble(prosoc_left = c(0, 1, 0, 1),
         condition   = c(0, 0, 1, 1)) %>% 
  expand(actor = c(str_c("new0", 1:9), str_c("new", 10:50)),
         nesting(prosoc_left, condition)) %>% 
  # this adds a row number, which will come in handy, later
  mutate(row = 1:n())

fitted(b12.4,
         newdata = nd,
         probs = c(.1, .9),
         allow_new_levels = T,
         sample_new_levels = "gaussian",
         summary = F,
         nsamples = 1) %>% 
  # wrangle
  as_tibble() %>% 
  set_names(1:200) %>% 
  gather() %>% 
  mutate(row = as.double(key)) %>% 
  # add the new data
  left_join(nd, by = "row") %>%
  # reformat `condition` for the plot
  mutate(condition = factor(str_c(prosoc_left, "/", condition), 
                            levels = c("0/0", "1/0", "0/1", "1/1"))) %>% 
  
  # plot!
  ggplot(aes(x = condition, y = value, group = actor)) +
  geom_line(alpha = 1/2, color = "blue") +
  ggtitle("50 simulated actors") +
  coord_cartesian(ylim = c(0, 1)) +
  # theme_fivethirtyeight() +
  theme(plot.title = element_text(size = 14, hjust = .5))
```


```{r, warning = F, message = F}
(
  f <-
  fitted(b12.4,
         newdata = nd,
         probs = c(.1, .9),
         allow_new_levels = T,
         sample_new_levels = "gaussian",
         summary = F,
         nsamples = 1) %>% 
  as_tibble() %>% 
  mutate(iter = 1:n_sim) %>% 
  gather(key, value, -iter) %>% 
  bind_cols(nd %>% 
              transmute(condition = factor(str_c(prosoc_left, "/", condition), 
                                           levels = c("0/0", "1/0", "0/1", "1/1"))) %>% 
              expand(condition, iter = 1:n_sim) %>% 
              select(-iter))
)
```



```{r, message = F}
# extract the posterior draws
post <- posterior_samples(b12.4)

# arrange the fixed effects (i.e., the population parameters)
fix_ef <-
  post %>% 
  # take only one posterior draw, the first row
  slice(1) %>%
  transmute(`0/0` = b_Intercept,
            `1/0` = b_Intercept + b_prosoc_left,
            `0/1` = b_Intercept,
            `1/1` = b_Intercept + b_prosoc_left + `b_prosoc_left:condition`) %>% 
  # duplicate the single posterior draw 1000 times
  uncount(weights = 1000)

# simulate 1000 unique random effects (actor-level intercept deviations)
set.seed(12.42)
ran_ef <-
  tibble(random_effect = rnorm(n = 1000, mean = 0, sd = post$sd_actor__Intercept))

# combine `ran_ef` and `fix_ef`
bind_cols(ran_ef, fix_ef) %>% 
  gather(-random_effect, key = "condition", value = "fixed_effect") %>% 
  mutate(condition = factor(condition, level = c("0/0", "1/0", "0/1", "1/1"))) %>%
  mutate(intercept = fixed_effect + random_effect) %>%
  mutate(prob = inv_logit_scaled(intercept)) %>%
  # make an iteration index
  mutate(iter = rep(1:1000, times = 4)) %>%
  # filter just the first 50
  filter(iter %in% c(1:50)) %>%
  
  # plot
  ggplot(aes(x = condition, y = prob, group = iter)) +
  geom_line(alpha = 1/2, color = "orange3") +
  ggtitle("50 simulated actors") +
  coord_cartesian(ylim = c(0, 1)) +
  # theme_fivethirtyeight() +
  theme(plot.title = element_text(size = 14, hjust = .5))


```


```{r, fig.width = 2.5, fig.height = 3.25}
p3 <-
  ran_and_fix_ef %>%
  mutate(iter = rep(1:1000, times = 4)) %>%
  filter(iter %in% c(1:50)) %>%
  
  ggplot(aes(x = condition, y = prob, group = iter)) +
  geom_line(alpha = 1/2, color = "orange3") +
  ggtitle("50 simulated actors") +
  coord_cartesian(ylim = c(0, 1)) +
  theme_fivethirtyeight() +
  theme(plot.title = element_text(size = 14, hjust = .5))

p3
```


## Session info {-}

```{r}
sessionInfo()
```

```{r, warning = F, echo = F}
rm()
```

```{r, echo = F, message = F, warning = F, results = "hide"}
ggplot2::theme_set(ggplot2::theme_grey())
bayesplot::color_scheme_set("blue")
pacman::p_unload(pacman::p_loaded(), character.only = TRUE)
```

# first 50

```{r}
# new data
nd <-
  tibble(prosoc_left = c(0, 1, 0, 1),
         condition   = c(0, 0, 1, 1)) %>% 
  expand(actor = str_c("new", 1:50),
         nesting(prosoc_left, condition)) %>% 
  # this adds a row number, which will come in handy, later
  mutate(row = 1:n())

# fitted
f <-
  fitted(b12.4,
         newdata = nd,
         allow_new_levels = T,
         sample_new_levels = "gaussian",
         summary = F,
         nsamples = 50)

# make a tibble and change the names to match the `row` values in `nd`
f %>% 
  as_tibble() %>% 
  set_names(1:200) %>% 
  # add an iteration index
  mutate(iter = 1:n()) %>% 
  # make it long
  gather(-iter, key = "row", value = "value") %>% 
  mutate(row = as.double(row)) %>% 
  # add the new data
  left_join(nd, by = "row") %>% 
  # extract the numbers from the names of the new actors
  mutate(actor_number = str_extract(actor, "\\d+") %>% as.double()) %>% 
  # only keep the posterior iterations that match the `actor_number` values
  filter(actor_number == iter) %>%
  # reformat `condition` for the plot
  mutate(condition = factor(str_c(prosoc_left, "/", condition), 
                            levels = c("0/0", "1/0", "0/1", "1/1"))) %>% 
  
  # plot!
  ggplot(aes(x = condition, y = value, group = actor)) +
  geom_line(alpha = 1/2, color = "blue") +
  ggtitle("50 simulated actors") +
  coord_cartesian(ylim = c(0, 1)) +
  # theme_fivethirtyeight() +
  theme(plot.title = element_text(size = 14, hjust = .5))
```



# full 16000
```{r}
f <-
  fitted(b12.4,
         newdata = nd,
         allow_new_levels = T,
         sample_new_levels = "gaussian",
         summary = F)

```

```{r}
# new data
nd <-
  tibble(prosoc_left = c(0, 1, 0, 1),
         condition   = c(0, 0, 1, 1)) %>% 
  expand(actor = c(str_c("new0", 1:9), str_c("new", 10:50)),
         nesting(prosoc_left, condition)) %>% 
  # this adds a row number, which will come in handy, later
  mutate(row = 1:n())

# fitted
f <-
  fitted(b12.4,
         newdata = nd,
         allow_new_levels = T,
         sample_new_levels = "gaussian",
         summary = F)

# this is terrible, but it works
set.seed(12.42)
f %>% 
  # make a tibble
  as_tibble() %>% 
  set_names(1:200) %>% 
  # add an iteration index
  mutate(iter = 1:n()) %>% 
  # make it long
  gather(-iter, key = "row", value = "value") %>% 
  mutate(row = as.double(row)) %>% 
  # add the new data
  left_join(nd, by = "row") %>% 
  # nest the data within `actor`
  nest(data = c(iter, row, value, prosoc_left, condition)) %>% 
  # for each level of `actor`, randomly select a number 
  # between 1 and 16,000 without replacement
  mutate(random_iter = sample(1:16000, size = n(), replace = F)) %>% 
  unnest(data) %>% 
  # now filter only the HMC iterations matching the random `random_iter` numbers
  filter(iter == random_iter) %>%
  # reformat `condition` for the plot
  mutate(condition = factor(str_c(prosoc_left, "/", condition), 
                            levels = c("0/0", "1/0", "0/1", "1/1"))) %>% 
  
  # plot!
  ggplot(aes(x = condition, y = value, group = actor)) +
  geom_line(alpha = 1/2, color = "blue") +
  ggtitle("50 simulated actors") +
  coord_cartesian(ylim = c(0, 1)) +
  # theme_fivethirtyeight() +
  theme(plot.title = element_text(size = 14, hjust = .5))
```




```{r}
nd <-
  tibble(prosoc_left = c(0, 1, 0, 1),
         condition   = c(0, 0, 1, 1)) %>% 
  expand(actor = c(str_c("new0", 1:9), str_c("new", 10:50)),
         nesting(prosoc_left, condition)) %>% 
  # this adds a row number, which will come in handy, later
  mutate(row = 1:n())

fitted(b12.4,
         newdata = nd,
         probs = c(.1, .9),
         allow_new_levels = T,
         sample_new_levels = "gaussian",
         summary = F,
         nsamples = 1) %>% 
  # wrangle
  as_tibble() %>% 
  set_names(1:200) %>% 
  gather() %>% 
  mutate(row = as.double(key)) %>% 
  # add the new data
  left_join(nd, by = "row") %>%
  # reformat `condition` for the plot
  mutate(condition = factor(str_c(prosoc_left, "/", condition), 
                            levels = c("0/0", "1/0", "0/1", "1/1"))) %>% 
  
  # plot!
  ggplot(aes(x = condition, y = value, group = actor)) +
  geom_line(alpha = 1/2, color = "blue") +
  ggtitle("50 simulated actors") +
  coord_cartesian(ylim = c(0, 1)) +
  # theme_fivethirtyeight() +
  theme(plot.title = element_text(size = 14, hjust = .5))
```








